### 2018-07-26 paper/paper.tex
Add very first version of paper

% !TeX document-id = {5ae73978-d7a8-4a15-a272-3b0597ff2241}
% This template has been tested with LLNCS DOCUMENT CLASS -- version 2.19 (04-Sep-2017)

% !TeX spellcheck = en-US
% !TeX encoding = utf8
% !TeX program = pdflatex
% !BIB program = bibtex
% -*- coding:utf-8 mod:LaTeX -*-

%"runningheads" enables:
%  - page number on page 2 onwards
%  - title/authors on even/odd pages
%This is good for other readers to enable proper archiving among other papers and pointing to
%content. Even if the title page states the title, when printed and stored in a folder, when
%blindly opening the folder, one could hit not the title page, but an arbitrary page. Therefore,
%it is good to have title printed on the pages, too.
\documentclass[runningheads,a4paper]{llncs}[2017/09/04]

%% If you need packages for other papers,
%% START COPYING HERE

% Set English as language and allow to write hyphenated"=words
%
% Even though `american`, `english` and `USenglish` are synonyms for babel package (according to https://tex.stackexchange.com/questions/12775/babel-english-american-usenglish), the llncs document class is prepared to avoid the overriding of certain names (such as "Abstract." -> "Abstract" or "Fig." -> "Figure") when using `english`, but not when using the other 2.
% english has to go last to set it as default language
\usepackage[ngerman,english]{babel}
%Hint by http://tex.stackexchange.com/a/321066/9075 -> enable "= as dashes
\addto\extrasenglish{\languageshorthands{ngerman}\useshorthands{"}}

\usepackage{ifluatex}
\ifluatex
  \usepackage{fontspec}
  \usepackage[english]{selnolig}
\fi

\iftrue % use default-font
  \ifluatex
    % use the better (sharper, ...) Latin Modern variant of Computer Modern
    \setmainfont{Latin Modern Roman}
    \setsansfont{Latin Modern Sans}
    \setmonofont{Latin Modern Mono} % "variable=false"
    %\setmonofont{Latin Modern Mono Prop} % "variable=true"
  \else
    % better font, similar to the default springer font
    % cfr-lm is preferred over lmodern. Reasoning at http://tex.stackexchange.com/a/247543/9075
    \usepackage[%
      rm={oldstyle=false,proportional=true},%
      sf={oldstyle=false,proportional=true},%
      tt={oldstyle=false,proportional=true,variable=false},%
      qt=false%
    ]{cfr-lm}
  \fi
\else
  % In case more space is needed, it is accepted to use Times New Roman
  \ifluatex
    \setmainfont{TeX Gyre Termes}
    \setsansfont[Scale=.9]{TeX Gyre Heros}
    % newtxtt looks good with times, but no equivalent for lualatex found,
    % therefore tried to replace with inconsolata.
    % However, inconsolata does not look good in the context of LNCS ...
    %\setmonofont[StylisticSet={1,3},Scale=.9]{inconsolata}
    % ... thus, we use the good old Latin Modern Mono font for source code.
    \setmonofont{Latin Modern Mono} % "variable=false"
    %\setmonofont{Latin Modern Mono Prop} % "variable=true"
  \else
    % overwrite cmodern with the Times variant
    \usepackage{newtxtext}
    \usepackage{newtxmath}
    \usepackage[zerostyle=b,scaled=.9]{newtxtt}
  \fi
\fi

\ifluatex
\else
  % fontenc and inputenc are not required when using lualatex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc} %support umlauts in the input
\fi

\usepackage{graphicx}

% backticks (`) are rendered as such in verbatim environment. See https://tex.stackexchange.com/a/341057/9075 for details.
\usepackage{upquote}

% Nicer tables (\toprule, \midrule, \bottomrule - see example)
\usepackage{booktabs}

%extended enumerate, such as \begin{compactenum}
\usepackage{paralist}

%put figures inside a text
%\usepackage{picins}
%use
%\piccaptioninside
%\piccaption{...}
%\parpic[r]{\includegraphics ...}
%Text...

% For easy quotations: \enquote{text}
% This package is very smart when nesting is applied, otherwise textcmds (see below) provides a shorter command
\usepackage{csquotes}

% For even easier quotations: \qq{text}
\usepackage{textcmds}

%enable margin kerning
\RequirePackage[%
  babel,%
  final,%
  expansion=alltext,%
  protrusion=alltext-nott]{microtype}%
% \texttt{test -- test} keeps the "--" as "--" (and does not convert it to an en dash)
\DisableLigatures{encoding = T1, family = tt* }

%tweak \url{...}
\usepackage{url}
%\urlstyle{same}
%improve wrapping of URLs - hint by http://tex.stackexchange.com/a/10419/9075
\makeatletter
\g@addto@macro{\UrlBreaks}{\UrlOrds}
\makeatother
%nicer // - solution by http://tex.stackexchange.com/a/98470/9075
%DO NOT ACTIVATE -> prevents line breaks
%\makeatletter
%\def\Url@twoslashes{\mathchar`\/\@ifnextchar/{\kern-.2em}{}}
%\g@addto@macro\UrlSpecials{\do\/{\Url@twoslashes}}
%\makeatother

% Diagonal lines in a table - http://tex.stackexchange.com/questions/17745/diagonal-lines-in-table-cell
% Slashbox is not available in texlive (due to licensing) and also gives bad results. This, we use diagbox
%\usepackage{diagbox}

% Required for package pdfcomment later
\usepackage{xcolor}

% For listings
\usepackage{listings}
\lstset{%
  basicstyle=\ttfamily,%
  columns=fixed,%
  basewidth=.5em,%
  xleftmargin=0.5cm,%
  captionpos=b}%
\renewcommand{\lstlistingname}{List.}
% Fix counter as described at https://tex.stackexchange.com/a/28334/9075
\usepackage{chngcntr}
\AtBeginDocument{\counterwithout{lstlisting}{section}}

% Enable nice comments
\usepackage{pdfcomment}
%
\newcommand{\commentontext}[2]{\colorbox{yellow!60}{#1}\pdfcomment[color={0.234 0.867 0.211},hoffset=-6pt,voffset=10pt,opacity=0.5]{#2}}
\newcommand{\commentatside}[1]{\pdfcomment[color={0.045 0.278 0.643},icon=Note]{#1}}
%
% Compatibality with packages todo, easy-todo, todonotes
\newcommand{\todo}[1]{\commentatside{#1}}
% Compatiblity with package fixmetodonotes
\newcommand{\TODO}[1]{\commentatside{#1}}

% Bibliopgraphy enhancements
%  - enable \cite[prenote][]{ref}
%  - enable \cite{ref1,ref2}
% Alternative: \usepackage{cite}, which enables \cite{ref1, ref2} only (otherwise: Error message: "White space in argument")

% Doc: http://texdoc.net/natbib
\usepackage[%
  square,        % for square brackets
  comma,         % use commas as separators
  numbers,       % for numerical citations;
%  sort,          % orders multiple citations into the sequence in which they appear in the list of references;
  sort&compress, % as sort but in addition multiple numerical citations
                 % are compressed if possible (as 3-6, 15);
]{natbib}
% In the bibliography, references have to be formatted as 1., 2., ... not [1], [2], ...
\renewcommand{\bibnumfmt}[1]{#1.}

\ifluatex
  % does not work when using luatex
  % see: https://tex.stackexchange.com/q/419288/9075
\else
  % Prepare more space-saving rendering of the bibliography
  % Source: https://tex.stackexchange.com/a/280936/9075
  \SetExpansion
  [ context = sloppy,
    stretch = 30,
    shrink = 60,
    step = 5 ]
  { encoding = {OT1,T1,TS1} }
  { }
\fi

% Put footnotes below floats
% Source: https://tex.stackexchange.com/a/32993/9075
\usepackage{stfloats}
\fnbelowfloat

% Enable that parameters of \cref{}, \ref{}, \cite{}, ... are linked so that a reader can click on the number an jump to the target in the document
\usepackage{hyperref}
% Enable hyperref without colors and without bookmarks
\hypersetup{hidelinks,
  colorlinks=true,
  allcolors=black,
  pdfstartview=Fit,
  breaklinks=true}
%
% Enable correct jumping to figures when referencing
\usepackage[all]{hypcap}

\usepackage[group-four-digits,per-mode=fraction]{siunitx}

%enable \cref{...} and \Cref{...} instead of \ref: Type of reference included in the link
\usepackage[capitalise,nameinlink]{cleveref}
%Nice formats for \cref
\crefname{section}{Sect.}{Sect.}
\Crefname{section}{Section}{Sections}
\crefname{listing}{\lstlistingname}{\lstlistingname}
\Crefname{listing}{Listing}{Listings}

%Intermediate solution for hyperlinked refs. See https://tex.stackexchange.com/q/132420/9075 for more information.
\newcommand{\Vlabel}[1]{\label[line]{#1}\hypertarget{#1}{}}
\newcommand{\lref}[1]{\hyperlink{#1}{\FancyVerbLineautorefname~\ref*{#1}}}

\usepackage{xspace}
%\newcommand{\eg}{e.\,g.\xspace}
%\newcommand{\ie}{i.\,e.\xspace}
\newcommand{\eg}{e.\,g.,\ }
\newcommand{\ie}{i.\,e.,\ }

%introduce \powerset - hint by http://matheplanet.com/matheplanet/nuke/html/viewtopic.php?topic=136492&post_id=997377
\DeclareFontFamily{U}{MnSymbolC}{}
\DeclareSymbolFont{MnSyC}{U}{MnSymbolC}{m}{n}
\DeclareFontShape{U}{MnSymbolC}{m}{n}{
  <-6>    MnSymbolC5
  <6-7>   MnSymbolC6
  <7-8>   MnSymbolC7
  <8-9>   MnSymbolC8
  <9-10>  MnSymbolC9
  <10-12> MnSymbolC10
  <12->   MnSymbolC12%
}{}
\DeclareMathSymbol{\powerset}{\mathord}{MnSyC}{180}

\ifluatex
\else
  % Enable copy and paste - also of numbers
  % This has to be done instead of \usepackage{cmap}, because it does not work together with cfr-lm.
  % See: https://tex.stackexchange.com/a/430599/9075
  \input glyphtounicode
  \pdfgentounicode=1
\fi

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%% END COPYING HERE

% Add copyright
% Do that for the final version or if you send it to colleagues
\iffalse
  %state: intended|submitted|llncs
  %you can add "crop" if the paper should be cropped to the format Springer is publishing
  \usepackage[intended]{llncsconf}

  \conference{name of the conference}

  %in case of "llncs" (final version!)
  %example: llncs{Anonymous et al. (eds). \emph{Proceedings of the International Conference on \LaTeX-Hacks}, LNCS~42. Some Publisher, 2016.}{0042}
  \llncs{book editors and title}{0042} %% 0042 is the start page
\fi

% For demonstration purposes only
\usepackage[math]{blindtext}
\usepackage{mwe}
\usepackage{pgfplots}

\begin{document}

\title{Modeling and Simulation of Load Balancing Strategies for Computing in High Energy Physics}
%If Title is too long, use \titlerunning
\titlerunning{Modeling and Simulation of Load Balancing Strategies}

%Single insitute
%\author{ \and Firstname Lastname}
%If there are too many authors, use \authorrunning
%\authorrunning{First Author et al.}
%\institute{Institute}

% Multiple insitutes - ALTERNATIVE to the above
 \author{
Ren\'e Caspart\inst{1} \and
Patrick Firnkes\inst{2} \and
Manuel Giffels\inst{1} \and
Anne Koziolek\inst{2} \and
G\"unter Quast\inst{1} \and
Ralf Reussner\inst{2} \and
Maximilian Stemmer-Grabow\inst{2}
 }

%
%If there are too many authors, use \authorrunning
\authorrunning{R. Caspart et al.}
%
\institute{
	Institute for Experimental Particle Physics (ETP) \and
	Institute for Program Structures and Data Organization (IPD) \\
	at Karlsruhe Institute of Technology (KIT)
}

\maketitle

\begin{abstract}
The amount of data to be processed by experiments in high energy physics (HEP) is tremendously increasing in the coming years. To cope with this increasing load, most efficient usage of the resources is mandatory. Furthermore, the computing resources for user jobs in HEP will be increasingly distributed and heterogeneous, resulting in a more difficult scheduling due the increasing complexity of the system. 
A first step to meet both challenges, a more efficient utilization of the grid and handling the rising complexity of the system, a simulation for the WLCG should be created. There is currently no simulation existing helping the operators of the grid to make the correct decisions while optimizing the load balancing strategy.

This paper shows a proof of concept in which the computing jobs at the Tier 1 center GridKa are modeled and simulated. To model the computing jobs we extended the Palladio simulator with a mechanism to simulate load balancing strategies. Furthermore, we implemented an automated model parameter analysis and model creation.

Finally, the simulation results are validated using real word performance data. Our results suggest that simulating larger parts of the grid is feasible and can help to optimize the utilization of the grid.
\end{abstract}

\begin{keywords}
  Performance Prediction,
  Modeling,
  Palladio,
  HEP,
  Load Balancing Strategies
\end{keywords}

\section{Introduction}
\label{sec:intro}

The amount of data processed by experiments in high energy physics (HEP) will drastically increase in the coming years due the upgrade of the LHC to the High-Luminosity Large Hadron Collider (HL-LHC), creating major challenges for the community of HEP \cite{community}.
\Cref{resources} shows the estimated CPU resources for the ATLAS experiment in the next ten years. As can be seen the flat computing budget model does not cover future needs for computing resource in the period from 2026 to 2028 due the start of the HL-LHC era. One of the key aspects to deal with this increasing load is to use the existing resources most efficiently.

Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources of user job. As cloud resources, like HNSciCloud and AWS, or institute clusters and Tier 3 centers will be used to compute them. This results in a more complex scheduling, because each resource has a different optimal job types and mixtures. Furthermore, this leads to a more dynamic data placement, where fewer centers hosts data and thus the remote access is increasing.

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{images/resources}
	\caption[]{Estimated CPU resources of ATLAS \cite{community}}
	\label{resources}
\end{figure}

To solve both challenges, a more efficient usage of the grid and handling the rising complexity of the system, a simulation of the WLCG should be created.
Currently, there is no model or simulation in existence to evaluate the load balancing strategies of the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
\Cref{related} describes different attempts to simulate the WLCG or parts of it. However, they showed scalability issues or the simulation does not obtain the necessary detail level to simulate effects on working nodes that required to evaluate load balancing strategies.

Optimizing load strategies without the help of simulations requires stable load, in order to preserve the comparability of the different strategies. This is neither possible to achieve for the WLCG nor for a single center like GridKa, because the nature of the load is highly dynamical. Another option would be to use an isolated test system for the evaluation. However, the system under inspection is too large and thus a similar test systems would not be financially affordable. A smaller test system could not deliver correct results because too much of the complexity would be taken away.
These reasons lead to the conclusion that a model and simulation of the performance of computing jobs is required to evaluate different load balancing strategies.

Our long term goal is to evaluate different load balancing strategies for the WLCG and additionally evaluate different design decisions affecting the performance of the grid. These gained insights will help the HEP community to solve the challenges caused by the HL-LHC upgrade.
As a proof of concept we modeled and simulated the performance of CMS computing jobs at the Tier 1 center GridKa.
The model and simulation will be created using the Palladio simulator. We extended this simulator to be able to simulate load balancing strategies. Furthermore, we created a automated model parameter analysis and model creation, to increase the usability.
This model enables us to evaluate strategies for GridKa and allows us to draw conclusions about the best optimizations of the load balancing for the WLCG. Furthermore, it can be used as a foundation for models and simulations of larger parts of the WLCG.

This paper is organized as follows: \Cref{sec:palladio} presents the Palladio simulator as a foundation of the simulation. \Cref{sec:process} gives a overview on the simulation process.\Cref{sec:model} describes how different elements of the computing jobs and the computing resources are represented in our model and the implementation of the necessarily extension of Palladio. \Cref{sec:param} explains how we determine the model parameters and the automated model creation. The validation is presented in \Cref{validation} and \Cref{related} puts our work in context with the related work. In \Cref{outlook} possible design decisions and how they could be modeled and simulated are presented. Finally, \Cref{conlusion} gives a conclusion on this paper.

\section{Foundations: Palladio}
\label{sec:palladio}
Palladio is a model driven software architecture simulator, which is implemented as an eclipse plugin and developed by Karlsruhe Institute of Technology (KIT), FZI Research Center for Information Technology, and University of Paderborn. Palladio predicts quality of software properties (e.g. performance) using several models of a system \cite{BECKER20093}.

These models are the component model, the assembly model, the resource model, the allocation model, and the usage model.
The component model specifies the structure and behavior of the components independently from their later usage. This allows reuse of the components, but requires its parametrization.
To represent how the components are connected to model the software architecture, the assembly model has to be created. 
The resource model describes the resource environment. It contains the number and characteristics of the resource containers on which the components could be deployed and the topology of the network connecting these containers.
Allocation models represent the mapping of the different components to the resource containers.
Finally, the usage model defines how a system is used regarding workload, user behavior, and parameters.
Once these models are created, Palladio can be used to simulate the system choosing one of several supported simulators.

Lehrig and Becker \cite{arch} extended Palladio with Architectural Templates to make it possible to model cloud computing environments efficiently.

Furthermore, Palladio's simulation approach SimuLizar was created to simulate self-adaptive systems \cite{becker2013simulizar}.
It was later extended to support the metrics scalability, elasticity, and efficiency \cite{arch}.

Palladio has been successfully used to solve several industrial problems and was used to optimize cloud infrastructure in the context of chemical computing as part of the CACTOS project \cite{rapidtesting}.

\section{Simulation Process}
\label{sec:process}
The starting point of the simulation process is the analysis of the job and working node data to extract performance parameters for the model. For example, these are the CPU demand of jobs, the frequency of job types or the CPU power of the nodes. This analysis is automated and described in \Cref{sec:param}. The next step is the generation of the model of GridKa and the injection of the previously extracted performance parameters. Both steps can be done via the UI of Palladio and do not require manually modeling. At this point the users have a valid and calibrated model of GridKa with its resource and the workload. The properties of this model are described in \Cref{sec:model}.

They can then either run the simulation or make changes to the model to simulate the effect of a design change. 
Finally, after running the simulation, they can inspect the results, such as core utilization of a node or throughput of jobs, using the integrated visualization tool or they can export the results and use a third party software to do the evaluation.

\section{Modeling Computing Jobs and Resources}
\label{sec:model}
This section describes how the computing jobs and working nodes are modeled using Palladio.
The component model describes each type of CMS computing job, such as analysis or reprocessing jobs, with their resource demands. This demands are the CPU demand, the I/O demand, the number of used CPU cores, the required job slots, and the number of processed events.

After a job is scheduled on a resource container it tries to acquire the specified number of job slots or waits if the node has not sufficient free slots. After that the CPU and I/O demand of the job are processed. To avoid one big I/O operation at the start of a job and make the processing of I/O more continuously, the processing is modeled using iterations, which represent the number of events. Thus, for each event there are I/O and CPU demands resources needed.
While processing, the job uses the specified number of cores to process the CPU demand. When the job finishes, it releases the job slots. All of these parameters are specified as a probability function and are extracted from measured performance data as described in \Cref{sec:param}.

Also each type of working node in GridKa is modeled in the resource model. Each node has a CPU and I/O processing rate, number of cores, number of job slots, and number of instances of this node type. These types of computing nodes are duplicated according to their specified number of instances as a preprocessing step of the simulation, resulting in the simulation of each working node.

At GridKa all computing jobs can run on all nodes, thus each job component is deployed on each computing node.

The usage model defines which and how many jobs are scheduled. To make sure that there is always a high load on the system, we used a closed workload with 5000 jobs. Each job type has a configurable share of the overall load.

We extended Palladio with a \textit{load balancing strategy action}, so that we can simulate load balancing strategies.
This load balancing strategy action is contained in a load balancing component.
In the assembly model all job components are connected with the load balancing component, which is the entry point of the system.
This action decides on which node a job should be scheduled based on the currently free job slots of each working node. The load balancing strategies can be extended to evaluate different strategies.

\section{Automated Model Parameter Analysis and Model Creation}
\label{sec:param}
\begin{itemize}
	\item data sources, matching, parameter extraction
	\item plugin to generate model
	\item cms share, only jobs that were run
	\item good usability
\end{itemize}

\section{Validation}
\label{validation}
To validate the model and simulation we simulated CMS jobs at GridKa in April 2018.
The job performance data was randomly separated into two equal sized calibration and test datasets.

The simulated throughput was higher than the measured one, because of that the we added a delay after each job, in order that the measured throughput matches the simulated one. The first factor for the higher throughput could be the missing overhead of pilot jobs. The second, more crucial, factor is the missing of jobs in the job data. We observed that 10-20\% of jobs are missing in the data, the cause for this is not known yet, however this should be investigated. 
After adding the delay the simulated throughput matches the measured one accurately, as \Cref{throughput} shows.
\Cref{shares} depicts the share of the simulated and measured job types, which also have a high accuracy.
\begin{figure}
	\centering
	\includegraphics[scale = 0.4]{images/durchsatz}
	\caption[]{Measured and Simulated Throughput}
	\label{throughput}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[scale = 0.4]{images/shares}
	\caption[]{Measured and Simulated Shares of Job Types}
	\label{shares}
\end{figure}

The simulated CPU efficiency is lower than the measured one, as shown in \Cref{utilization}. The reason for this is likely missing jobs in the job data. Because of that is the calibrated throughput is too low resulting in a lower utilization.

Simulating one week of CMS jobs at GridKa takes about 18 minutes on a laptop. Thus, it would feasible to simulate several Tiers at once in a sufficient time. However, for simulating larger parts of the WLCG a more macroscopic model, which focuses on the interaction of the Tiers and does not model their inner structure in detail,
is recommended.

\begin{figure}
	\vspace{-0.5ex}
	\centering
		\begin{tikzpicture}[scale = 1]
		\begin{axis}[
		xlabel={Time / s},
		ylabel={Total CPU Time / Walltime},
		ymin=0, ymax=1,
		xmin=0, xmax=604800,
		legend cell align={left},
		]
		\addplot[color=red] coordinates {
			(0, 0.61)			
			(604800, 0.61)
		};
		\addplot[color=blue] coordinates {	
			(14000,0.420457263335)
			(15000,0.455205081375)
			(16000,0.450933749429)
			(17000,0.46853252847)
			(18000,0.474029527868)
			(19000,0.482875361196)
			(20000,0.472149795914)
			(21000,0.448813851682)
			(22000,0.430706201872)
			(23000,0.401599999586)
			(24000,0.376755783861)
			(25000,0.38749824392)
			(26000,0.422990841095)
			(27000,0.425760711131)
			(28000,0.43904605249)
			(29000,0.437695380408)
			(30000,0.433314619731)
			(31000,0.448449713155)
			(32000,0.441020436611)
			(33000,0.444429125129)
			(34000,0.461008578928)
			(35000,0.470284257719)
			(36000,0.447847515795)
			(37000,0.441187264182)
			(38000,0.433373025865)
			(39000,0.426778346)
			(40000,0.440334830361)
			(41000,0.452643935542)
			(42000,0.452142136739)
			(43000,0.452471374183)
			(44000,0.445302566248)
			(45000,0.437814297445)
			(46000,0.445663557997)
			(47000,0.451677101561)
			(48000,0.442604461856)
			(49000,0.453882226309)
			(50000,0.444581524102)
			(51000,0.443425493542)
			(52000,0.424588550506)
			(53000,0.424918434251)
			(54000,0.438754707788)
			(55000,0.431586617405)
			(56000,0.432664270787)
			(57000,0.431460959005)
			(58000,0.43776201488)
			(59000,0.452017801875)
			(60000,0.445027944703)
			(61000,0.447754439714)
			(62000,0.441042254855)
			(63000,0.448994039435)
			(64000,0.447949217961)
			(65000,0.440335544034)
			(66000,0.432872781216)
			(67000,0.444765956656)
			(68000,0.445503771932)
			(69000,0.435313479888)
			(70000,0.440723673441)
			(71000,0.453581869677)
			(72000,0.455011550524)
			(73000,0.445251024271)
			(74000,0.461605049304)
			(75000,0.470460077846)
			(76000,0.467995526805)
			(77000,0.495748712572)
			(78000,0.49681048222)
			(79000,0.495635118813)
			(80000,0.476519088357)
			(81000,0.469046301151)
			(82000,0.459260808853)
			(83000,0.443186601975)
			(84000,0.453927784554)
			(85000,0.460996975902)
			(86000,0.439846472086)
			(87000,0.42853879571)
			(88000,0.425940455445)
			(89000,0.418188095172)
			(90000,0.441124921832)
			(91000,0.430042869457)
			(92000,0.454088618748)
			(93000,0.447812967768)
			(94000,0.472828979798)
			(95000,0.503483458242)
			(96000,0.49269529217)
			(97000,0.485731894357)
			(98000,0.500299990804)
			(99000,0.489833797954)
			(100000,0.486536917579)
			(101000,0.470900103725)
			(102000,0.473487499528)
			(103000,0.472852199897)
			(104000,0.465907927377)
			(105000,0.46843843884)
			(106000,0.469393804824)
			(107000,0.448337382107)
			(108000,0.453458598472)
			(109000,0.477909863511)
			(110000,0.484074360628)
			(111000,0.471446789672)
			(112000,0.483311476858)
			(113000,0.492794346306)
			(114000,0.494774125893)
			(115000,0.478760296453)
			(116000,0.463289501118)
			(117000,0.470560118374)
			(118000,0.46261989666)
			(119000,0.448321324595)
			(120000,0.4391456597)
			(121000,0.44616088099)
			(122000,0.443067693493)
			(123000,0.443765615997)
			(124000,0.45489288544)
			(125000,0.457761907159)
			(126000,0.459890981374)
			(127000,0.457265336864)
			(128000,0.450661574709)
			(129000,0.456319671876)
			(130000,0.455222292845)
			(131000,0.444168136077)
			(132000,0.455292020739)
			(133000,0.459739659432)
			(134000,0.474266033943)
			(135000,0.476719078979)
			(136000,0.454265351086)
			(137000,0.432403127216)
			(138000,0.426941224365)
			(139000,0.435303571358)
			(140000,0.445594945267)
			(141000,0.457945319262)
			(142000,0.467041094326)
			(143000,0.471275880951)
			(144000,0.448803750886)
			(145000,0.46057752759)
			(146000,0.450222072137)
			(147000,0.447748079045)
			(148000,0.463476880757)
			(149000,0.457547341501)
			(150000,0.464860371781)
			(151000,0.458939834833)
			(152000,0.449435321435)
			(153000,0.445809287248)
			(154000,0.452098810214)
			(155000,0.436323981134)
			(156000,0.435163726955)
			(157000,0.430055068196)
			(158000,0.432586012365)
			(159000,0.437297849322)
			(160000,0.424896762305)
			(161000,0.433031154001)
			(162000,0.42905370657)
			(163000,0.412624449914)
			(164000,0.412491612196)
			(165000,0.399704545922)
			(166000,0.407748045902)
			(167000,0.409904586648)
			(168000,0.412048703271)
			(169000,0.422106417772)
			(170000,0.431756334979)
			(171000,0.432552532485)
			(172000,0.461650235913)
			(173000,0.450819280269)
			(174000,0.455592609364)
			(175000,0.461503577581)
			(176000,0.463421686809)
			(177000,0.456182986217)
			(178000,0.465859886059)
			(179000,0.469016662103)
			(180000,0.457242602407)
			(181000,0.438524914279)
			(182000,0.419623185788)
			(183000,0.434867384803)
			(184000,0.436354031073)
			(185000,0.431955984805)
			(186000,0.426743598615)
			(187000,0.433639371186)
			(188000,0.447367993742)
			(189000,0.441424080277)
			(190000,0.460353123518)
			(191000,0.460813912595)
			(192000,0.455819448956)
			(193000,0.467462336883)
			(194000,0.475473651551)
			(195000,0.464487515787)
			(196000,0.45335938384)
			(197000,0.435968082782)
			(198000,0.459905581295)
			(199000,0.443261087286)
			(200000,0.438900909062)
			(201000,0.470264321796)
			(202000,0.451214857632)
			(203000,0.468171752558)
			(204000,0.459209586602)
			(205000,0.459254298111)
			(206000,0.47101680367)
			(207000,0.463033170114)
			(208000,0.4712737853)
			(209000,0.465450336139)
			(210000,0.455801602051)
			(211000,0.461908888488)
			(212000,0.458205085034)
			(213000,0.461468054298)
			(214000,0.453173854006)
			(215000,0.454871547866)
			(216000,0.444018420001)
			(217000,0.460956695722)
			(218000,0.465661865182)
			(219000,0.465518305154)
			(220000,0.4711530613)
			(221000,0.459010429226)
			(222000,0.464906045582)
			(223000,0.453027764206)
			(224000,0.455831376049)
			(225000,0.469807813442)
			(226000,0.456897450438)
			(227000,0.446722893031)
			(228000,0.444441981548)
			(229000,0.431491369988)
			(230000,0.448834255688)
			(231000,0.445076241297)
			(232000,0.435089055194)
			(233000,0.437592817834)
			(234000,0.437875351009)
			(235000,0.434588735814)
			(236000,0.44507195818)
			(237000,0.439254948215)
			(238000,0.444706733582)
			(239000,0.447474042265)
			(240000,0.446329940764)
			(241000,0.455749636545)
			(242000,0.419361606614)
			(243000,0.445397279873)
			(244000,0.452348733449)
			(245000,0.434373580372)
			(246000,0.43809435824)
			(247000,0.457201403905)
			(248000,0.458802551256)
			(249000,0.440383453863)
			(250000,0.445017392448)
			(251000,0.450084408166)
			(252000,0.441726249534)
			(253000,0.43194266247)
			(254000,0.445568321745)
			(255000,0.436278081802)
			(256000,0.439570299584)
			(257000,0.445136501732)
			(258000,0.45903255575)
			(259000,0.450554481769)
			(260000,0.446171160379)
			(261000,0.458690240049)
			(262000,0.470220369345)
			(263000,0.480684091886)
			(264000,0.456854673542)
			(265000,0.450233371349)
			(266000,0.457899571788)
			(267000,0.453334820977)
			(268000,0.460295743577)
			(269000,0.457349902193)
			(270000,0.462039377481)
			(271000,0.481382348042)
			(272000,0.491382702182)
			(273000,0.484220311688)
			(274000,0.491652039182)
			(275000,0.486379984979)
			(276000,0.494978853484)
			(277000,0.490824451822)
			(278000,0.461792687666)
			(279000,0.469822396762)
			(280000,0.48576001703)
			(281000,0.481488945743)
			(282000,0.480743747455)
			(283000,0.486090522042)
			(284000,0.487356854396)
			(285000,0.465745257924)
			(286000,0.472330225888)
			(287000,0.492995112907)
			(288000,0.489510571313)
			(289000,0.493418264292)
			(290000,0.489755865611)
			(291000,0.489256145131)
			(292000,0.483996042192)
			(293000,0.471077668813)
			(294000,0.471398859559)
			(295000,0.461127782716)
			(296000,0.458754727805)
			(297000,0.451312728894)
			(298000,0.457666739683)
			(299000,0.463072024695)
			(300000,0.452352488238)
			(301000,0.432056039316)
			(302000,0.442534543126)
			(303000,0.464020942755)
			(304000,0.465788968846)
			(305000,0.475194751884)
			(306000,0.460429089784)
			(307000,0.440203075328)
			(308000,0.440392507843)
			(309000,0.469031285375)
			(310000,0.457737564909)
			(311000,0.450884070739)
			(312000,0.467259163074)
			(313000,0.467018061082)
			(314000,0.466325864949)
			(315000,0.4515369024)
			(316000,0.451628960618)
			(317000,0.449715736896)
			(318000,0.453599972995)
			(319000,0.46598650055)
			(320000,0.458071641171)
			(321000,0.459795875349)
			(322000,0.463776572273)
			(323000,0.441893234874)
			(324000,0.450990407649)
			(325000,0.469838695448)
			(326000,0.454971976955)
			(327000,0.458423340158)
			(328000,0.455021879641)
			(329000,0.474962772272)
			(330000,0.468921434665)
			(331000,0.463586286578)
			(332000,0.472336338416)
			(333000,0.480203877337)
			(334000,0.499976146642)
			(335000,0.499528976694)
			(336000,0.486582556042)
			(337000,0.486946662392)
			(338000,0.472474569197)
			(339000,0.47914245784)
			(340000,0.475286986461)
			(341000,0.458659120511)
			(342000,0.463977117388)
			(343000,0.48570346922)
			(344000,0.47347942112)
			(345000,0.450392499597)
			(346000,0.445409073982)
			(347000,0.454343590968)
			(348000,0.446283787024)
			(349000,0.441991195931)
			(350000,0.457080478112)
			(351000,0.475143781352)
			(352000,0.453825073455)
			(353000,0.466696828065)
			(354000,0.461398654993)
			(355000,0.476432737168)
			(356000,0.484493658032)
			(357000,0.478381386037)
			(358000,0.48791767455)
			(359000,0.471135835878)
			(360000,0.4510006391)
			(361000,0.459254358243)
			(362000,0.45940296355)
			(363000,0.455276346902)
			(364000,0.460372976975)
			(365000,0.46432780643)
			(366000,0.468703252372)
			(367000,0.460690627792)
			(368000,0.456596843643)
			(369000,0.453188747838)
			(370000,0.444026572421)
			(371000,0.469540885626)
			(372000,0.477111174668)
			(373000,0.453778831255)
			(374000,0.451432024952)
			(375000,0.467745860758)
			(376000,0.488699820134)
			(377000,0.47828379946)
			(378000,0.470780021427)
			(379000,0.460858540737)
			(380000,0.468116574189)
			(381000,0.476739105023)
			(382000,0.465550165705)
			(383000,0.466704565487)
			(384000,0.460951873868)
			(385000,0.474085333555)
			(386000,0.483774319139)
			(387000,0.450257163363)
			(388000,0.432558156082)
			(389000,0.441761962984)
			(390000,0.471357238783)
			(391000,0.476402662876)
			(392000,0.478773054016)
			(393000,0.479343982261)
			(394000,0.476955005539)
			(395000,0.464686525062)
			(396000,0.465867162002)
			(397000,0.473713210956)
			(398000,0.455977533182)
			(399000,0.453556194091)
			(400000,0.468032168961)
			(401000,0.473372878696)
			(402000,0.460103363611)
			(403000,0.450437880682)
			(404000,0.444189180728)
			(405000,0.440631553941)
			(406000,0.440814465656)
			(407000,0.439132208745)
			(408000,0.443680589194)
			(409000,0.439364305134)
			(410000,0.445754620988)
			(411000,0.459675216489)
			(412000,0.434411774646)
			(413000,0.460859998054)
			(414000,0.470068659819)
			(415000,0.46398628613)
			(416000,0.46092602178)
			(417000,0.470952398074)
			(418000,0.464918768054)
			(419000,0.457250642935)
			(420000,0.452696999147)
			(421000,0.443597876602)
			(422000,0.442688896448)
			(423000,0.446975959689)
			(424000,0.459962628411)
			(425000,0.458639133149)
			(426000,0.45869214177)
			(427000,0.459339918847)
			(428000,0.469954533509)
			(429000,0.466801073197)
			(430000,0.479585876798)
			(431000,0.473889297072)
			(432000,0.468945113898)
			(433000,0.46819782837)
			(434000,0.460713604747)
			(435000,0.467085533437)
			(436000,0.464877942532)
			(437000,0.441101434716)
			(438000,0.435732064901)
			(439000,0.448217580287)
			(440000,0.447917450562)
			(441000,0.443407621096)
			(442000,0.443456810485)
			(443000,0.438987312155)
			(444000,0.467286398042)
			(445000,0.469189496526)
			(446000,0.467728960254)
			(447000,0.469175237549)
			(448000,0.469335300568)
			(449000,0.463820736067)
			(450000,0.458428492443)
			(451000,0.448441949348)
			(452000,0.453900123964)
			(453000,0.459953017199)
			(454000,0.448381040252)
			(455000,0.465700695725)
			(456000,0.451531261356)
			(457000,0.456563899123)
			(458000,0.45720539038)
			(459000,0.451801077935)
			(460000,0.464657248779)
			(461000,0.459373104135)
			(462000,0.469958682024)
			(463000,0.467903777714)
			(464000,0.445428785915)
			(465000,0.442499436959)
			(466000,0.447134373326)
			(467000,0.438282409091)
			(468000,0.43839540036)
			(469000,0.419945925767)
			(470000,0.418452547982)
			(471000,0.438307200322)
			(472000,0.45350446074)
			(473000,0.468786686762)
			(474000,0.456254084131)
			(475000,0.453297591133)
			(476000,0.469078411526)
			(477000,0.457354866601)
			(478000,0.476430954364)
			(479000,0.475029005373)
			(480000,0.479985350112)
			(481000,0.475381159607)
			(482000,0.461543267652)
			(483000,0.451592900155)
			(484000,0.45260612709)
			(485000,0.458008617252)
			(486000,0.444827616864)
			(487000,0.447400533469)
			(488000,0.452399703406)
			(489000,0.451208289224)
			(490000,0.469255926366)
			(491000,0.469056554934)
			(492000,0.448427595423)
			(493000,0.460791545132)
			(494000,0.465807144119)
			(495000,0.459466948041)
			(496000,0.457398936667)
			(497000,0.454270854364)
			(498000,0.44030861198)
			(499000,0.441478571167)
			(500000,0.448416540804)
			(501000,0.447708055398)
			(502000,0.44394661343)
			(503000,0.437604010563)
			(504000,0.43279917301)
			(505000,0.448202463548)
			(506000,0.450335081656)
			(507000,0.456367109826)
			(508000,0.476797822515)
			(509000,0.464004036098)
			(510000,0.447462462599)
			(511000,0.460128351384)
			(512000,0.456610531662)
			(513000,0.461027700674)
			(514000,0.452647871516)
			(515000,0.472383445652)
			(516000,0.480048107691)
			(517000,0.449495011852)
			(518000,0.424921363583)
			(519000,0.430957570675)
			(520000,0.42946792577)
			(521000,0.435679220854)
			(522000,0.454543429655)
			(523000,0.449873483822)
			(524000,0.441368528175)
			(525000,0.434726970329)
			(526000,0.446558845708)
			(527000,0.436523566317)
			(528000,0.436567752145)
			(529000,0.440939357596)
			(530000,0.442182493949)
			(531000,0.446092107838)
			(532000,0.451187916749)
			(533000,0.439896893747)
			(534000,0.449011137767)
			(535000,0.444120664286)
			(536000,0.426232253223)
			(537000,0.438918286378)
			(538000,0.44181502269)
			(539000,0.432995321972)
			(540000,0.434879809332)
			(541000,0.427145288337)
			(542000,0.427872289185)
			(543000,0.417701942996)
			(544000,0.419515873503)
			(545000,0.439790579048)
			(546000,0.43762292406)
			(547000,0.443042538052)
			(548000,0.455727191114)
			(549000,0.447614026634)
			(550000,0.456624751487)
			(551000,0.452793939028)
			(552000,0.440726166681)
			(553000,0.449038613212)
			(554000,0.468536381832)
			(555000,0.444002288289)
			(556000,0.452092079245)
			(557000,0.456883014276)
			(558000,0.453112451125)
			(559000,0.441037791264)
			(560000,0.442347996342)
			(561000,0.436264841204)
			(562000,0.438959132682)
			(563000,0.4451724287)
			(564000,0.458215756554)
			(565000,0.448690251696)
			(566000,0.453197454907)
			(567000,0.453794387302)
			(568000,0.444846550094)
			(569000,0.449982902679)
			(570000,0.447267180536)
			(571000,0.445828406211)
			(572000,0.446909040851)
			(573000,0.453980157655)
			(574000,0.457292498947)
			(575000,0.448977443253)
			(576000,0.446078471261)
			(577000,0.434934699022)
			(578000,0.45949539644)
			(579000,0.477069726029)
			(580000,0.478983629734)
			(581000,0.459161976806)
			(582000,0.470964357654)
			(583000,0.466734817389)
			(584000,0.451817556376)
			(585000,0.438500345769)
			(586000,0.435540269671)
			(587000,0.436722886274)
			(588000,0.452696266979)
			(589000,0.444922652517)
			(590000,0.414546052368)
			(591000,0.40666087591)
			(592000,0.431870620823)
			(593000,0.456065081956)
			(594000,0.45959389509)
			(595000,0.445709712809)
			(596000,0.455310403564)
			(597000,0.474126321375)
			(598000,0.466945015295)
			(599000,0.442916390259)
			(600000,0.463327317157)
			(601000,0.463538228748)
			(602000,0.484414072165)
			(603000,0.470969958228)
			(604000,0.357975119833)

		};
		\addlegendentry{Measured (average)};
		\addlegendentry{Simulated};
		\end{axis}
		\end{tikzpicture}
		\caption[]{Measured and Simulated CPU Efficiency}
		\label{utilization}
\end{figure}

\section{Related Work}
\label{related}
The MONARC project was started 1998 to help with the development of the initial design of the WLCG at CERN \cite{monarc2000models}.
However, after they initial design was developed, the project was discontinued in 2000. 
In 2004 a second version of the MONARC simulator was released with several improvements, such as simulation of data replication. However, MONARC2 showed scalability issues when simulating a Tier-2 site and accuracy issues regarding network load \cite{1742-6596-331-7-072038}.

There are a variety of grid and cloud simulators existing, of which SimGrid and CloudSim are two popular ones. 
However, both of them are missing graphical model editors which makes the models of the computing jobs are time-consuming to create \cite{simgrid}\cite{cloudsim}. The support of graphical editors is a big advantage of Palladio and increases its usability and efficiency.

CACTOS is an approach to cloud infrastructure automation and optimization \cite{cactos}.
One part of the approach is the CACTOS Runtime Toolkit for monitoring and resource management and another part is the CACTOS Prediction Toolkit for evaluation of alternative data center deployment scenarios and resource management algorithms. The Prediction Toolkit is built on Palladio and SimuLizar.
A first case studies in the context of chemical computing showed that the Prediction Toolkit has a high accuracy, but it was not tested on large systems yet \cite{rapidtesting}. However, due the required Runtime Toolkit CACTOS was not applicable in our use case.

Korenkov et. al simulated the JINR Tier1 site in 2016 \cite{jinr}. They focused on the effects of the location of the required job data the throughput of the site. However, they neither modeled each working node nor modeled details like job slots, thus this simulation would not be usable to simulate load balancing strategies on working node level.

Hushchyn et al. created a simulation for the LHCb Grid in 2017 using SimGrid \cite{lhcb}. This simulation focused on which site a job should be placed based on data location. They modeled each site coarse granularly, thus it is not possible to simulate scheduling effects inside a computing center. In contrast our model is more detailed and is able to simulate effects regarding scheduling on working node level.

\section{Outlook}
\label{outlook}
The next steps are the evaluation of different design decisions using our simulation. In this section a few design decisions are described that could be simulated.

One possible design decision is a more sophisticated load balancing strategy that aims for a better resource utilization of the working nodes by scheduling a balanced share of CPU and IO heavy jobs to it. To simulate this a new scheduling strategy must be implemented.

Another design decision would be to experiment with different sizes of pilot jobs, assuming that smaller pilot jobs result in a better utilization due less wasted CPU capacity in the draining phase of a pilot. For this use case the model must be extended with the explicit modeling of pilots jobs.

Additionally, the effects of integrating cloud resources could be simulated. The model must be extended with the cloud resources and the required network load to load the files from a site. Then a good scheduling strategy for integration of cloud resources could be evaluated, like only scheduling CPU heavy jobs to cloud resources and keep I/O heavy jobs near the data.

Also the effects of data caches could be simulated. Firstly, several sites must be implemented to simulate their interactions. Secondly, the jobs must be parametrized with the possibility of remote reading. Then the data cache should be modeled. Finally, the performance improvements of the data cache by transferring data from the cache and not from sites can be simulated.
\section{Conclusion}
\label{conlusion}
We successfully simulated the CMS computing jobs at GridKa. To enable easy usage of the simulation we implemented an automated model parameter extraction and model creation. Furthermore, we extended Palladio to be able to simulate load balancing strategies. Our first results suggest that this approach could deliver accurate results regarding CPU efficiency, however it must be investigated what the reason for the data inconsistency and missing job information is. Regarding scalability we are confident that several computing centers could be simulated in such a detailed manner like we did with the CMS computing jobs at GridKa. For use cases that require the simulation of larger parts of the WLCG we suggest a more macroscopic model that focuses on the interaction of the sites and models the sites internals on a less detailed level.

This proof of concept showed that an automated model parameter extraction and model creation is possible and that Palladio can be used to evaluate load balancing strategies for computing in HEP. In contrast to the existing simulations our simulation models a site in a more detailed way, enabling the evaluation of load balancing strategies on working node level. Furthermore, we are confident that our approach helps to improve utilization of the grid. The next step is the evaluation of a design decision mentioned in \Cref{outlook}.

\renewcommand{\bibsection}{\section*{References}} % requried for natbib to have "References" printed and as section*, not chapter*
% Use natbib compatbile splncsnat style.
% It does provide all features of splncs03, but is developed in a clean way.
% Source: http://phaseportrait.blogspot.de/2011/02/natbib-compatible-bibtex-style-bst-file.html
\bibliographystyle{splncsnat}
\begingroup
  \ifluatex
    %try to activate if bibliography looks ugly
    %\sloppy
  \else
    \microtypecontext{expansion=sloppy}
  \fi
  \small % ensure correct font size for the bibliography
  \bibliography{paper}
\endgroup

% Enfore empty line after bibliography
\ \\
%

\end{document}

### 2018-08-14 paper/paper.tex
Update paper
Use latex template and add impreovements from meeting.

@@
-% !TeX document-id = {5ae73978-d7a8-4a15-a272-3b0597ff2241}
-% This template has been tested with LLNCS DOCUMENT CLASS -- version 2.19 (04-Sep-2017)
-
-% !TeX spellcheck = en-US
-% !TeX encoding = utf8
-% !TeX program = pdflatex
-% !BIB program = bibtex
-% -*- coding:utf-8 mod:LaTeX -*-
-
-%"runningheads" enables:
-%  - page number on page 2 onwards
-%  - title/authors on even/odd pages
-%This is good for other readers to enable proper archiving among other papers and pointing to
-%content. Even if the title page states the title, when printed and stored in a folder, when
-%blindly opening the folder, one could hit not the title page, but an arbitrary page. Therefore,
-%it is good to have title printed on the pages, too.
-\documentclass[runningheads,a4paper]{llncs}[2017/09/04]
-
-%% If you need packages for other papers,
-%% START COPYING HERE
-
-% Set English as language and allow to write hyphenated"=words
-%
-% Even though `american`, `english` and `USenglish` are synonyms for babel package (according to https://tex.stackexchange.com/questions/12775/babel-english-american-usenglish), the llncs document class is prepared to avoid the overriding of certain names (such as "Abstract." -> "Abstract" or "Fig." -> "Figure") when using `english`, but not when using the other 2.
-% english has to go last to set it as default language
-\usepackage[ngerman,english]{babel}
-%Hint by http://tex.stackexchange.com/a/321066/9075 -> enable "= as dashes
-\addto\extrasenglish{\languageshorthands{ngerman}\useshorthands{"}}
-
-\usepackage{ifluatex}
-\ifluatex
-  \usepackage{fontspec}
-  \usepackage[english]{selnolig}
-\fi
-
-\iftrue % use default-font
-  \ifluatex
-    % use the better (sharper, ...) Latin Modern variant of Computer Modern
-    \setmainfont{Latin Modern Roman}
-    \setsansfont{Latin Modern Sans}
-    \setmonofont{Latin Modern Mono} % "variable=false"
-    %\setmonofont{Latin Modern Mono Prop} % "variable=true"
-  \else
-    % better font, similar to the default springer font
-    % cfr-lm is preferred over lmodern. Reasoning at http://tex.stackexchange.com/a/247543/9075
-    \usepackage[%
-      rm={oldstyle=false,proportional=true},%
-      sf={oldstyle=false,proportional=true},%
-      tt={oldstyle=false,proportional=true,variable=false},%
-      qt=false%
-    ]{cfr-lm}
-  \fi
-\else
-  % In case more space is needed, it is accepted to use Times New Roman
-  \ifluatex
-    \setmainfont{TeX Gyre Termes}
-    \setsansfont[Scale=.9]{TeX Gyre Heros}
-    % newtxtt looks good with times, but no equivalent for lualatex found,
-    % therefore tried to replace with inconsolata.
-    % However, inconsolata does not look good in the context of LNCS ...
-    %\setmonofont[StylisticSet={1,3},Scale=.9]{inconsolata}
-    % ... thus, we use the good old Latin Modern Mono font for source code.
-    \setmonofont{Latin Modern Mono} % "variable=false"
-    %\setmonofont{Latin Modern Mono Prop} % "variable=true"
-  \else
-    % overwrite cmodern with the Times variant
-    \usepackage{newtxtext}
-    \usepackage{newtxmath}
-    \usepackage[zerostyle=b,scaled=.9]{newtxtt}
-  \fi
-\fi
-
-\ifluatex
-\else
-  % fontenc and inputenc are not required when using lualatex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc} %support umlauts in the input
-\fi
+\documentclass[a4paper]{jpconf}
 
@@
 
-% backticks (`) are rendered as such in verbatim environment. See https://tex.stackexchange.com/a/341057/9075 for details.
-\usepackage{upquote}
-
-% Nicer tables (\toprule, \midrule, \bottomrule - see example)
-\usepackage{booktabs}
-
-%extended enumerate, such as \begin{compactenum}
-\usepackage{paralist}
-
-%put figures inside a text
-%\usepackage{picins}
-%use
-%\piccaptioninside
-%\piccaption{...}
-%\parpic[r]{\includegraphics ...}
-%Text...
-
-% For easy quotations: \enquote{text}
-% This package is very smart when nesting is applied, otherwise textcmds (see below) provides a shorter command
-\usepackage{csquotes}
-
-% For even easier quotations: \qq{text}
-\usepackage{textcmds}
-
-%enable margin kerning
-\RequirePackage[%
-  babel,%
-  final,%
-  expansion=alltext,%
-  protrusion=alltext-nott]{microtype}%
-% \texttt{test -- test} keeps the "--" as "--" (and does not convert it to an en dash)
-\DisableLigatures{encoding = T1, family = tt* }
-
-%tweak \url{...}
-\usepackage{url}
-%\urlstyle{same}
-%improve wrapping of URLs - hint by http://tex.stackexchange.com/a/10419/9075
-\makeatletter
-\g@addto@macro{\UrlBreaks}{\UrlOrds}
-\makeatother
-%nicer // - solution by http://tex.stackexchange.com/a/98470/9075
-%DO NOT ACTIVATE -> prevents line breaks
-%\makeatletter
-%\def\Url@twoslashes{\mathchar`\/\@ifnextchar/{\kern-.2em}{}}
-%\g@addto@macro\UrlSpecials{\do\/{\Url@twoslashes}}
-%\makeatother
-
-% Diagonal lines in a table - http://tex.stackexchange.com/questions/17745/diagonal-lines-in-table-cell
-% Slashbox is not available in texlive (due to licensing) and also gives bad results. This, we use diagbox
-%\usepackage{diagbox}
-
-% Required for package pdfcomment later
-\usepackage{xcolor}
-
-% For listings
-\usepackage{listings}
-\lstset{%
-  basicstyle=\ttfamily,%
-  columns=fixed,%
-  basewidth=.5em,%
-  xleftmargin=0.5cm,%
-  captionpos=b}%
-\renewcommand{\lstlistingname}{List.}
-% Fix counter as described at https://tex.stackexchange.com/a/28334/9075
-\usepackage{chngcntr}
-\AtBeginDocument{\counterwithout{lstlisting}{section}}
-
-% Enable nice comments
-\usepackage{pdfcomment}
-%
-\newcommand{\commentontext}[2]{\colorbox{yellow!60}{#1}\pdfcomment[color={0.234 0.867 0.211},hoffset=-6pt,voffset=10pt,opacity=0.5]{#2}}
-\newcommand{\commentatside}[1]{\pdfcomment[color={0.045 0.278 0.643},icon=Note]{#1}}
-%
-% Compatibality with packages todo, easy-todo, todonotes
-\newcommand{\todo}[1]{\commentatside{#1}}
-% Compatiblity with package fixmetodonotes
 \newcommand{\TODO}[1]{\commentatside{#1}}
-
-% Bibliopgraphy enhancements
-%  - enable \cite[prenote][]{ref}
-%  - enable \cite{ref1,ref2}
-% Alternative: \usepackage{cite}, which enables \cite{ref1, ref2} only (otherwise: Error message: "White space in argument")
-
-% Doc: http://texdoc.net/natbib
-\usepackage[%
-  square,        % for square brackets
-  comma,         % use commas as separators
-  numbers,       % for numerical citations;
-%  sort,          % orders multiple citations into the sequence in which they appear in the list of references;
-  sort&compress, % as sort but in addition multiple numerical citations
-                 % are compressed if possible (as 3-6, 15);
-]{natbib}
-% In the bibliography, references have to be formatted as 1., 2., ... not [1], [2], ...
-\renewcommand{\bibnumfmt}[1]{#1.}
-
-\ifluatex
-  % does not work when using luatex
-  % see: https://tex.stackexchange.com/q/419288/9075
-\else
-  % Prepare more space-saving rendering of the bibliography
-  % Source: https://tex.stackexchange.com/a/280936/9075
-  \SetExpansion
-  [ context = sloppy,
-    stretch = 30,
-    shrink = 60,
-    step = 5 ]
-  { encoding = {OT1,T1,TS1} }
-  { }
-\fi
-
-% Put footnotes below floats
-% Source: https://tex.stackexchange.com/a/32993/9075
-\usepackage{stfloats}
-\fnbelowfloat
-
-% Enable that parameters of \cref{}, \ref{}, \cite{}, ... are linked so that a reader can click on the number an jump to the target in the document
-\usepackage{hyperref}
-% Enable hyperref without colors and without bookmarks
-\hypersetup{hidelinks,
-  colorlinks=true,
-  allcolors=black,
-  pdfstartview=Fit,
-  breaklinks=true}
-%
-% Enable correct jumping to figures when referencing
-\usepackage[all]{hypcap}
-
-\usepackage[group-four-digits,per-mode=fraction]{siunitx}
-
-%enable \cref{...} and \Cref{...} instead of \ref: Type of reference included in the link
 \usepackage[capitalise,nameinlink]{cleveref}
@@
 \Crefname{listing}{Listing}{Listings}
-
-%Intermediate solution for hyperlinked refs. See https://tex.stackexchange.com/q/132420/9075 for more information.
-\newcommand{\Vlabel}[1]{\label[line]{#1}\hypertarget{#1}{}}
-\newcommand{\lref}[1]{\hyperlink{#1}{\FancyVerbLineautorefname~\ref*{#1}}}
-
-\usepackage{xspace}
-%\newcommand{\eg}{e.\,g.\xspace}
-%\newcommand{\ie}{i.\,e.\xspace}
-\newcommand{\eg}{e.\,g.,\ }
-\newcommand{\ie}{i.\,e.,\ }
-
-%introduce \powerset - hint by http://matheplanet.com/matheplanet/nuke/html/viewtopic.php?topic=136492&post_id=997377
-\DeclareFontFamily{U}{MnSymbolC}{}
-\DeclareSymbolFont{MnSyC}{U}{MnSymbolC}{m}{n}
-\DeclareFontShape{U}{MnSymbolC}{m}{n}{
-  <-6>    MnSymbolC5
-  <6-7>   MnSymbolC6
-  <7-8>   MnSymbolC7
-  <8-9>   MnSymbolC8
-  <9-10>  MnSymbolC9
-  <10-12> MnSymbolC10
-  <12->   MnSymbolC12%
-}{}
-\DeclareMathSymbol{\powerset}{\mathord}{MnSyC}{180}
-
-\ifluatex
-\else
-  % Enable copy and paste - also of numbers
-  % This has to be done instead of \usepackage{cmap}, because it does not work together with cfr-lm.
-  % See: https://tex.stackexchange.com/a/430599/9075
-  \input glyphtounicode
-  \pdfgentounicode=1
-\fi
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-%% END COPYING HERE
-
-% Add copyright
-% Do that for the final version or if you send it to colleagues
-\iffalse
-  %state: intended|submitted|llncs
-  %you can add "crop" if the paper should be cropped to the format Springer is publishing
-  \usepackage[intended]{llncsconf}
-
-  \conference{name of the conference}
-
-  %in case of "llncs" (final version!)
-  %example: llncs{Anonymous et al. (eds). \emph{Proceedings of the International Conference on \LaTeX-Hacks}, LNCS~42. Some Publisher, 2016.}{0042}
-  \llncs{book editors and title}{0042} %% 0042 is the start page
-\fi
-
-% For demonstration purposes only
-\usepackage[math]{blindtext}
-\usepackage{mwe}
 \usepackage{pgfplots}
 
+\bibliographystyle{iopart-num}
+
 \begin{document}
-
 \title{Modeling and Simulation of Load Balancing Strategies for Computing in High Energy Physics}
-%If Title is too long, use \titlerunning
-\titlerunning{Modeling and Simulation of Load Balancing Strategies}
-
-%Single insitute
-%\author{ \and Firstname Lastname}
-%If there are too many authors, use \authorrunning
-%\authorrunning{First Author et al.}
-%\institute{Institute}
-
-% Multiple insitutes - ALTERNATIVE to the above
- \author{
-Ren\'e Caspart\inst{1} \and
-Patrick Firnkes\inst{2} \and
-Manuel Giffels\inst{1} \and
-Anne Koziolek\inst{2} \and
-G\"unter Quast\inst{1} \and
-Ralf Reussner\inst{2} \and
-Maximilian Stemmer-Grabow\inst{2}
- }
-
-%
-%If there are too many authors, use \authorrunning
-\authorrunning{R. Caspart et al.}
-%
-\institute{
-	Institute for Experimental Particle Physics (ETP) \and
-	Institute for Program Structures and Data Organization (IPD) \\
-	at Karlsruhe Institute of Technology (KIT)
-}
-
-\maketitle
+
+\author{Ren\'e Caspart${}^{1}$,
+	Patrick Firnkes${}^{2}$,
+	Manuel Giffels${}^{1}$,
+	Anne Koziolek${}^{2}$,
+	G\"unter Quast${}^{1}$,
+	Ralf Reussner${}^{2}$ and 
+	Maximilian Stemmer-Grabow${}^{2}$}
+
+\address{
+	${}^{1}$Institute for Experimental Particle Physics (ETP) \\
+	${}^{2}$Institute for Program Structures and Data Organization (IPD) \\
+	at Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany}
+
+\ead{koziolek@kit.edu}
 
@@
 Furthermore, the computing resources for user jobs in HEP will be increasingly distributed and heterogeneous, resulting in a more difficult scheduling due the increasing complexity of the system.
-A first step to meet both challenges, a more efficient utilization of the grid and handling the rising complexity of the system, a simulation for the WLCG should be created.
+	A first step to meet both challenges, a more efficient utilization of the grid and coping with the rising complexity of the system, a simulation for the WLCG should be created.
 There is currently no simulation existing helping the operators of the grid to make the correct decisions while optimizing the load balancing strategy.
@@
 
-\begin{keywords}
-  Performance Prediction,
-  Modeling,
-  Palladio,
-  HEP,
-  Load Balancing Strategies
-\end{keywords}
-
 \section{Introduction}
@@
 The amount of data processed by experiments in high energy physics (HEP) will drastically increase in the coming years due the upgrade of the LHC to the High-Luminosity Large Hadron Collider (HL-LHC), creating major challenges for the community of HEP \cite{community}.
-\Cref{resources} shows the estimated CPU resources for the ATLAS experiment in the next ten years.
-As can be seen the flat computing budget model does not cover future needs for computing resource in the period from 2026 to 2028 due the start of the HL-LHC era.
-One of the key aspects to deal with this increasing load is to use the existing resources most efficiently.
-
-Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources of user job.
-As cloud resources, like HNSciCloud and AWS, or institute clusters and Tier 3 centers will be used to compute them.
+\Cref{resources} shows the estimated required and provided CPU resources for the ATLAS experiment in the next ten years.
+Assuming a flat computing budget model, the estimated provided resources do not cover future needs for computing resource in the period from 2026 to 2028 due the start of the HL-LHC era.
+One of the key aspects to deal with this increasing load is the more efficient usage of existing resources.
+
+Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources of user jobs.
+Cloud resources, like HNSciCloud and AWS, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs.
 This results in a more complex scheduling, because each resource has a different optimal job types and mixtures.
@@
 
-To solve both challenges, a more efficient usage of the grid and handling the rising complexity of the system, a simulation of the WLCG should be created.
+To solve both challenges, a more efficient usage of the grid and coping with the rising complexity of the system, a simulation of the WLCG should be created.
 Currently, there is no model or simulation in existence to evaluate the load balancing strategies of the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
 \Cref{related} describes different attempts to simulate the WLCG or parts of it.
-However, they showed scalability issues or the simulation does not obtain the necessary detail level to simulate effects on working nodes that required to evaluate load balancing strategies.
-
-Optimizing load strategies without the help of simulations requires stable load, in order to preserve the comparability of the different strategies.
+However, they showed scalability issues or the simulations do not have the necessary detail level to simulate effects on working nodes that is necessary to evaluate load balancing strategies.
+
+Optimizing load balancing strategies without the help of simulations requires stable load, in order to preserve the comparability of the different strategies.
 This is neither possible to achieve for the WLCG nor for a single center like GridKa, because the nature of the load is highly dynamical.
@@
 However, the system under inspection is too large and thus a similar test systems would not be financially affordable.
-A smaller test system could not deliver correct results because too much of the complexity would be taken away.
+A smaller test system could not deliver correct results because too much of the complexity of the system would be taken away.
 These reasons lead to the conclusion that a model and simulation of the performance of computing jobs is required to evaluate different load balancing strategies.
@@
 These gained insights will help the HEP community to solve the challenges caused by the HL-LHC upgrade.
-As a proof of concept we modeled and simulated the performance of CMS computing jobs at the Tier 1 center GridKa.
-The model and simulation will be created using the Palladio simulator.
+
+The contribution of this paper is the modeling and simulating the performance of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
+The model and simulation are created using the Palladio simulator.
 We extended this simulator to be able to simulate load balancing strategies.
 Furthermore, we created a automated model parameter analysis and model creation, to increase the usability.
-This model enables us to evaluate strategies for GridKa and allows us to draw conclusions about the best optimizations of the load balancing for the WLCG.
+This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG.
 Furthermore, it can be used as a foundation for models and simulations of larger parts of the WLCG.
@@
 This paper is organized as follows: \Cref{sec:palladio} presents the Palladio simulator as a foundation of the simulation.
-\Cref{sec:process} gives a overview on the simulation process.\Cref{sec:model} describes how different elements of the computing jobs and the computing resources are represented in our model and the implementation of the necessarily extension of Palladio.
+\Cref{sec:process} gives a overview on the simulation process.
+\Cref{sec:model} describes how different elements of the computing jobs and the computing resources are represented in our model and the implementation of the necessarily extension of Palladio.
 \Cref{sec:param} explains how we determine the model parameters and the automated model creation.
@@
 These models are the component model, the assembly model, the resource model, the allocation model, and the usage model.
-The component model specifies the structure and behavior of the components independently from their later usage. This allows reuse of the components, but requires its parametrization.
-To represent how the components are connected to model the software architecture, the assembly model has to be created. 
-The resource model describes the resource environment.
+\begin{itemize}
+	\item The component model: Specifies the structure and behavior of the components independently from their later usage. This allows reuse of the components, but requires its parametrization.
+	\item The assembly model: Represents how the components are connected to model the software architecture.
+	\item The resource model: Describes the resource environment.
 It contains the number and characteristics of the resource containers on which the components could be deployed and the topology of the network connecting these containers.
-Allocation models represent the mapping of the different components to the resource containers.
-Finally, the usage model defines how a system is used regarding workload, user behavior, and parameters.
+	\item The allocation model: Represents the mapping of the different components to the resource containers.
+	\item The usage model: Defines the usage of a system regarding workload, user behavior, and parameters.
+
+\end{itemize}
 Once these models are created, Palladio can be used to simulate the system choosing one of several supported simulators.
@@
 Lehrig and Becker \cite{arch} extended Palladio with Architectural Templates to make it possible to model cloud computing environments efficiently.
-
 Furthermore, Palladio's simulation approach SimuLizar was created to simulate self-adaptive systems \cite{becker2013simulizar}.
@@
 \label{sec:process}
-The starting point of the simulation process is the analysis of the job and working node data to extract performance parameters for the model.
-For example, these are the CPU demand of jobs, the frequency of job types or the CPU power of the nodes.
-This analysis is automated and described in \Cref{sec:param}.
+
+\Cref{process} shows the simulation process starting with the matching of the data sets to receiving a valid model of the site and its jobs.
+
+The starting point of the simulation process is to match the data sets from different data sources as a prerequisite  of the analysis.
+In our proof of concept we have three data sources: The CERN WMArchive, CMS GridKa job data and GridKa node performance data.
+
+After matching and grouping the data for job types the analysis of the job and working node data to extract performance parameters for the model can take place.
+For example, these are the CPU demand of jobs, the frequency of the job types or the CPU power of the nodes.
+This analysis and the data set matching are automated and described in \Cref{sec:param}.
+
 The next step is the generation of the model of GridKa and the injection of the previously extracted performance parameters.
@@
 Finally, after running the simulation, they can inspect the results, such as core utilization of a node or throughput of jobs, using the integrated visualization tool or they can export the results and use a third party software to do the evaluation.
+
+\begin{figure}
+	\centering
+	\includegraphics[scale = 0.4]{images/process}
+	\caption[]{The Simulation Process}
+	\label{process}
+\end{figure}
 
@@
 
-After a job is scheduled on a resource container it tries to acquire the specified number of job slots or waits if the node has not sufficient free slots.
+After a job is scheduled on a resource container it acquires the specified number of job slots or waits if the node has not sufficient free slots.
 After that the CPU and I/O demand of the job are processed.
-To avoid one big I/O operation at the start of a job and make the processing of I/O more continuously, the processing is modeled using iterations, which represent the number of events.
-Thus, for each event there are I/O and CPU demands resources needed.
+To avoid one big I/O operation at the start of a job and make the processing of I/O more continuously, the processing is modeled using iterations that reflect the number of events.
+Thus, for each iteration there are I/O and CPU resources needed.
+Rather than simulating each event, we simulate a part of the processed events, because the simulation of each event would harm the scalability of the simulation.
 While processing, the job uses the specified number of cores to process the CPU demand.
-When the job finishes, it releases the job slots.
+To simulate not explicitly modeled elements of the jobs such as the overhead created by pilot jobs we added a delay to each job.
+This delay is calibrated in order to match the measured throughput.
+After the job finished, it releases its acquired job slots.
 All of these parameters are specified as a probability function and are extracted from measured performance data as described in \Cref{sec:param}.
 
-Also each type of working node in GridKa is modeled in the resource model.
-Each node has a CPU and I/O processing rate, number of cores, number of job slots, and number of instances of this node type.
-These types of computing nodes are duplicated according to their specified number of instances as a preprocessing step of the simulation, resulting in the simulation of each working node.
-
-At GridKa all computing jobs can run on all nodes, thus each job component is deployed on each computing node.
+Each type of working node in GridKa is modeled in the resource model.
+Each node type has a CPU and I/O processing rate, number of cores, number of job slots, and number of instances of this node type.
+The types of computing nodes are duplicated according to their specified number of instances as a preprocessing step of the simulation, resulting in the simulation of each working node.
+
+At GridKa every type of computing job can run on every node, thus each job type is allocated on each computing node in the allocation model.
 
 The usage model defines which and how many jobs are scheduled.
-To make sure that there is always a high load on the system, we used a closed workload with 5000 jobs.
+To make sure that there is always a high load on the system, we used a closed workload with enough jobs that the system is fully utilized.
 Each job type has a configurable share of the overall load.
@@
 
-The simulated throughput was higher than the measured one, because of that the we added a delay after each job, in order that the measured throughput matches the simulated one.
-The first factor for the higher throughput could be the missing overhead of pilot jobs.
-The second, more crucial, factor is the missing of jobs in the job data.
-We observed that 10-20\% of jobs are missing in the data, the cause for this is not known yet, however this should be investigated.
-After adding the delay the simulated throughput matches the measured one accurately, as \Cref{throughput} shows.
+As mentioned in \Cref{sec:model} a delay was added to the job to simulate not explicitly modeled elements.
+This delay was calibrated so that the measured throughput matches the simulated one of the calibration dataset.
+The simulated throughput of the test dataset matches the measured one accurately with only 1\% difference.
 \Cref{shares} depicts the share of the simulated and measured job types, which also have a high accuracy.
-\begin{figure}
-	\centering
-	\includegraphics[scale = 0.4]{images/durchsatz}
-	\caption[]{Measured and Simulated Throughput}
-	\label{throughput}
-\end{figure}
+
 \begin{figure}
@@
 The simulated CPU efficiency is lower than the measured one, as shown in \Cref{utilization}.
-The reason for this is likely missing jobs in the job data.
-Because of that is the calibrated throughput is too low resulting in a lower utilization.
+The reason for this are likely missing jobs in the job data.
+We observed that a significant amount of data is missing, the cause for this is not known yet, however this should be investigated.
+Because of that the calibrated throughput is too low, resulting in a lower utilization.
+Another reason is the missing utilization created by pilot jobs that are not explicitly modeled yet.
 
@@
 The MONARC project was started 1998 to help with the development of the initial design of the WLCG at CERN \cite{monarc2000models}.
-However, after they initial design was developed, the project was discontinued in 2000. 
+However, after the initial design was developed, the project was discontinued in 2000 and thus no model and simulation of the WLCG was created. 
 In 2004 a second version of the MONARC simulator was released with several improvements, such as simulation of data replication.
@@
 There are a variety of grid and cloud simulators existing, of which SimGrid and CloudSim are two popular ones. 
-However, both of them are missing graphical model editors which makes the models of the computing jobs are time-consuming to create \cite{simgrid}\cite{cloudsim}.
+However, both of them are missing graphical model editors which makes the modeling of the computing jobs time-consuming \cite{simgrid}\cite{cloudsim}.
 The support of graphical editors is a big advantage of Palladio and increases its usability and efficiency.
+ parameterize
 
@@
 Korenkov et.
-al simulated the JINR Tier1 site in 2016 \cite{jinr}.
-They focused on the effects of the location of the required job data the throughput of the site.
-However, they neither modeled each working node nor modeled details like job slots, thus this simulation would not be usable to simulate load balancing strategies on working node level.
-
-Hushchyn et al.
+al simulated the JINR Tier1 site in 2016 \cite{jinr} and Hushchyn et al.
 created a simulation for the LHCb Grid in 2017 using SimGrid \cite{lhcb}.
-This simulation focused on which site a job should be placed based on data location.
-They modeled each site coarse granularly, thus it is not possible to simulate scheduling effects inside a computing center.
-In contrast our model is more detailed and is able to simulate effects regarding scheduling on working node level.
+Both simulations focused on the effects of the location of the required job data on the throughput of the sites.
+
+However, they neither modeled each working node nor modeled details of the scheduling process like job slots, thus both simulations would not be usable to simulate load balancing strategies on working node level.
 
@@
 \label{outlook}
-The next steps are the evaluation of different design decisions using our simulation. In this section a few design decisions are described that could be simulated.
+Our next steps are the evaluation of different design decisions. In this section a few design decisions are described that could be simulated.
 
@@
 Additionally, the effects of integrating cloud resources could be simulated.
-The model must be extended with the cloud resources and the required network load to load the files from a site.
-Then a good scheduling strategy for integration of cloud resources could be evaluated, like only scheduling CPU heavy jobs to cloud resources and keep I/O heavy jobs near the data.
+The model must be extended with the cloud resources and the required network load to access the files from a site.
+Then a good load balancing strategy for the  integration of cloud resources could be evaluated, like only scheduling CPU heavy jobs to cloud resources and keeping I/O heavy jobs near the data.
 
 Also the effects of data caches could be simulated.
-Firstly, several sites must be implemented to simulate their interactions.
-Secondly, the jobs must be parametrized with the possibility of remote reading.
+Firstly, several sites must be modeled to simulate their interactions.
+Secondly, the jobs must be parametrized with the possibility of remote access.
 Then the data cache should be modeled.
@@
 Our first results suggest that this approach could deliver accurate results regarding CPU efficiency, however it must be investigated what the reason for the data inconsistency and missing job information is.
-Regarding scalability we are confident that several computing centers could be simulated in such a detailed manner like we did with the CMS computing jobs at GridKa.
+Regarding scalability we are confident that several computing centers could be simulated on a detailed level.
 For use cases that require the simulation of larger parts of the WLCG we suggest a more macroscopic model that focuses on the interaction of the sites and models the sites internals on a less detailed level.
@@
 In contrast to the existing simulations our simulation models a site in a more detailed way, enabling the evaluation of load balancing strategies on working node level.
-Furthermore, we are confident that our approach helps to improve utilization of the grid.
+Furthermore, we are confident that our approach helps to improve utilization of the grid and thus helps solving the challenges created by the HL-LHC update.
 The next step is the evaluation of a design decision mentioned in \Cref{outlook}.
 
-\renewcommand{\bibsection}{\section*{References}} % requried for natbib to have "References" printed and as section*, not chapter*
-% Use natbib compatbile splncsnat style.
-% It does provide all features of splncs03, but is developed in a clean way.
-% Source: http://phaseportrait.blogspot.de/2011/02/natbib-compatible-bibtex-style-bst-file.html
-\bibliographystyle{splncsnat}
-\begingroup
-  \ifluatex
-    %try to activate if bibliography looks ugly
-    %\sloppy
-  \else
-    \microtypecontext{expansion=sloppy}
-  \fi
-  \small % ensure correct font size for the bibliography
+\section*{References}
 \bibliography{paper}
-\endgroup
-
-% Enfore empty line after bibliography
-\ \\
-%
 

### 2018-08-17 paper/paper.tex
Add todo
Next iteration of paper

@@
 
-\newcommand{\TODO}[1]{\commentatside{#1}}
+\usepackage{todonotes}
 \usepackage[capitalise,nameinlink]{cleveref}
@@
 	G\"unter Quast${}^{1}$,
-	Ralf Reussner${}^{2}$ and 
+	Ralf Reussner${}^{2}$, 
 	Maximilian Stemmer-Grabow${}^{2}$}
@@
 Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources of user jobs.
-Cloud resources, like HNSciCloud and AWS, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs.
-This results in a more complex scheduling, because each resource has a different optimal job types and mixtures.
-Furthermore, this leads to a more dynamic data placement, where fewer centers hosts data and thus the remote access is increasing.
+Cloud resources, such as HNSciCloud and AWS, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs.
+This results in a more complex scheduling, because each resource has different optimal job types and mixtures.
+Furthermore, this leads to a more dynamic data placement, where fewer centers host data and thus remote access is increasing.
 
@@
 
-To solve both challenges, a more efficient usage of the grid and coping with the rising complexity of the system, a simulation of the WLCG should be created.
+To solve both challenges, a more efficient usage of the grid and coping with the rising complexity of the system, a simulation for the WLCG should be created.
 Currently, there is no model or simulation in existence to evaluate the load balancing strategies of the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
@@
 
-The contribution of this paper is the modeling and simulating the performance of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
+The contribution of this paper is the modeling and simulating of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
 The model and simulation are created using the Palladio simulator.
@@
 \Cref{sec:process} gives a overview on the simulation process.
-\Cref{sec:model} describes how different elements of the computing jobs and the computing resources are represented in our model and the implementation of the necessarily extension of Palladio.
-\Cref{sec:param} explains how we determine the model parameters and the automated model creation.
+\Cref{sec:model} describes how different elements of the computing jobs and the computing resources are represented in our model.
+\Cref{sec:param} explains how we determine the model parameters and the automated model creation works.
 The validation is presented in \Cref{validation} and \Cref{related} puts our work in context with the related work.
@@
 Furthermore, Palladio's simulation approach SimuLizar was created to simulate self-adaptive systems \cite{becker2013simulizar}.
-It was later extended to support the metrics scalability, elasticity, and efficiency \cite{arch}.
+It was later extended to support the metrics scalability, elasticity, and efficiency.
 
@@
 
-\Cref{process} shows the simulation process starting with the matching of the data sets to receiving a valid model of the site and its jobs.
+\Cref{process} depicts the simulation process starting with matching of the data sets to receiving a valid model of the site and its jobs.
 
-The starting point of the simulation process is to match the data sets from different data sources as a prerequisite  of the analysis.
+The starting point of the simulation process is matching the data sets from different data sources as a prerequisite  of the analysis.
 In our proof of concept we have three data sources: The CERN WMArchive, CMS GridKa job data and GridKa node performance data.
+These data sets have to be matched because each set contains different information about an executed job, but non of them contains the complete information about it.
 
-After matching and grouping the data for job types the analysis of the job and working node data to extract performance parameters for the model can take place.
+After matching and grouping the data for job types, the analysis of the job and working node data to extract performance parameters for the model takes place.
 For example, these are the CPU demand of jobs, the frequency of the job types or the CPU power of the nodes.
@@
 After a job is scheduled on a resource container it acquires the specified number of job slots or waits if the node has not sufficient free slots.
-After that the CPU and I/O demand of the job are processed.
-To avoid one big I/O operation at the start of a job and make the processing of I/O more continuously, the processing is modeled using iterations that reflect the number of events.
+Afterwards the CPU and I/O demands of the job are processed.
+To avoid one big I/O operation at the start of a job and to make the processing of I/O more continuously, the processing is modeled using iterations that reflect the number of events.
 Thus, for each iteration there are I/O and CPU resources needed.
-Rather than simulating each event, we simulate a part of the processed events, because the simulation of each event would harm the scalability of the simulation.
+Rather than simulating each event, we simulate a fraction of the processed events, because the simulation of each event would harm the scalability of the simulation.
 While processing, the job uses the specified number of cores to process the CPU demand.
@@
 
-As mentioned in \Cref{sec:model} a delay was added to the job to simulate not explicitly modeled elements.
+As mentioned in \Cref{sec:model} a delay was added to the job to simulate the effects of not explicitly modeled elements.
 This delay was calibrated so that the measured throughput matches the simulated one of the calibration dataset.
@@
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
-	\includegraphics[scale = 0.4]{images/shares}
+	\includegraphics[scale = 0.6]{images/shares}
 	\caption[]{Measured and Simulated Shares of Job Types}
@@
 There are a variety of grid and cloud simulators existing, of which SimGrid and CloudSim are two popular ones. 
-However, both of them are missing graphical model editors which makes the modeling of the computing jobs time-consuming \cite{simgrid}\cite{cloudsim}.
+However, both of them are implemented as libraries and not as standalone simulators.
+Thus, both are missing graphical model editors which makes the modeling of the computing jobs time-consuming \cite{simgrid}\cite{cloudsim}.
 The support of graphical editors is a big advantage of Palladio and increases its usability and efficiency.
- parameterize
+Another advantage of Palladio is the separation of the system into different models, which enables one to change on part of the system without modifying others.
+Furthermore, both do lack the support of an integrated visualization tool to display the results.
 
@@
 al simulated the JINR Tier1 site in 2016 \cite{jinr} and Hushchyn et al.
-created a simulation for the LHCb Grid in 2017 using SimGrid \cite{lhcb}.
+created a simulation for the LHCb Grid in 2017 \cite{lhcb}.
 Both simulations focused on the effects of the location of the required job data on the throughput of the sites.
-
 However, they neither modeled each working node nor modeled details of the scheduling process like job slots, thus both simulations would not be usable to simulate load balancing strategies on working node level.
@@
 \label{outlook}
-Our next steps are the evaluation of different design decisions. In this section a few design decisions are described that could be simulated.
+Our next steps are the evaluation of different design decisions. In this section a selection of possible design decisions that could be simulated are presented.
 
@@
 \label{conlusion}
+
+Computing resources for user analyses will become increasingly distributed and heterogeneous in the future and 
+scheduling will be a key component for the efficient usage of the grid.
+We are confident that our approach helps to improve utilization of the grid and to be able to cope with the rising complexity of the system, thus serves the HEP community as a tool to solve the challenges created by the HL-LHC update.
+
 We successfully simulated the CMS computing jobs at GridKa.
@@
 In contrast to the existing simulations our simulation models a site in a more detailed way, enabling the evaluation of load balancing strategies on working node level.
-Furthermore, we are confident that our approach helps to improve utilization of the grid and thus helps solving the challenges created by the HL-LHC update.
-The next step is the evaluation of a design decision mentioned in \Cref{outlook}.
-
+The next step is the evaluation of design decisions mentioned in \Cref{outlook}.
 \section*{References}

### 2018-09-16 paper/paper.tex
Fix citation
Replace simulation process figure with pdf version
Re-add section about parameter extraction

@@
 	\centering
-	\includegraphics[scale = 0.4]{images/process}
+	\includegraphics[width=0.6\linewidth]{images/process}
 	\caption[]{The Simulation Process}
@@
 
-\section{Automated Model Parameter Analysis and Model Creation}
+\section{Model Parameter Extraction and Model Creation}
 \label{sec:param}
+
+To obtain a comprehensive view both of the resource usage of CMS jobs executed at a WLCG grid site as well as the resources available there to process them, combining information from multiple data sources is required:
 \begin{itemize}
-	\item data sources, matching, parameter extraction
-	\item plugin to generate model
-	\item cms share, only jobs that were run
-	\item good usability
+    \item We acquire Historical job monitoring information from several global CMS monitoring systems: Job reports from JobMonitoring and Workflow and Data Management framework job reports (FWJR documents) are archived on the \emph{analytix} Hadoop cluster.
+    The latter are being archived on the cluster as part of the WMArchive project \cite{WMArchiveWorkloadArchive2018}.
+    We collect job information using the \emph{CMSSpark} framework by Kuznetsov et al. \cite{cmsspark}.
+
+    For this proof of concept, only jobs executed at the GridKa site were considered, but these data sources can be used to collect job reports from arbitrary WLCG sites.
+    As these data sets contain different metrics for each executed job and not every job is contained in each data set, they are matched to each other to identify records that describe the same job execution prior to analysis.
+    \item Site-specific performance data sets include the available nodes at the site and their performance information (including node benchmarks such as HS06 scores), which are available from local site monitoring systems.
+    The simulation currently only includes computing jobs from CMS.
+    As GridKa also processes jobs for other experiments, the number of available nodes at the site is scaled down to match the share of CMS workloads at the site in the time frame used for calibration.
+    This share of resources is also collected from site-local monitoring information, as well as CPU efficiency reference data for the workloads executed in the same time frame.
 \end{itemize}
+To allow quick extraction of model parameters from these data sets, we implemented data set matching as well as the following analysis steps in an automated, configurable model parameter extraction tool written in Python.
+Matching job information to local node performance information allows their resource usage to be decoupled from the hardware used to execute them, e.g.
+by combining the CPU time they used with node benchmarks of the machine they were executed on.
+
+For the purposes of this proof of concept, jobs were then grouped according to their type as indicated in the job reports, but the extraction framework allows to be configured for arbitrary groupings of the extracted jobs.
+Job resource requirement distributions for the different job types are extracted as distributions based on the empirical distributions in the data sets used for calibration, in the form of discretized probability density functions.
+
+The parameter extraction tool exports model parameter sets as JSON files.
+To simplify calibrating simulation models with these parameter sets, we created a Palladio plugin that can be used to automatically complete a blueprint of the simulation model (which defines the structure of the model as described in section \ref{sec:model}) with arbitrary parameter sets.
+This ensures an efficient and flexible model creation process.
 
@@
 However, both of them are implemented as libraries and not as standalone simulators.
-Thus, both are missing graphical model editors which makes the modeling of the computing jobs time-consuming \cite{simgrid}\cite{cloudsim}.
+Thus, both are missing graphical model editors which makes the modeling of the computing jobs time-consuming \cite{simgrid,cloudsim}.
 The support of graphical editors is a big advantage of Palladio and increases its usability and efficiency.

### 2018-10-02 paper/paper.tex
Add suggestions from review

@@
 	\centering
-	\includegraphics[width=1.0\linewidth]{images/resources}
+	\includegraphics[width=0.7\linewidth]{images/resources}
 	\caption[]{Estimated CPU resources of ATLAS \cite{community}}
@@
 The contribution of this paper is the modeling and simulating of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
-The model and simulation are created using the Palladio simulator.
-We extended this simulator to be able to simulate load balancing strategies.
-Furthermore, we created a automated model parameter analysis and model creation, to increase the usability.
+The model and simulation are created using the Palladio simulator, see \Cref{sec:palladio} for details on Palladio.
+This leads to the research question which modifications are needed so that Palladio can be used to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio.
+To realise it we created an extension for the simulator that enables us to simulate load balancing strategies and furthermore, we created an automated model parameter analysis and model creation, in order to increase the usability of the simulation.
 This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG.
@@
 \label{sec:process}
-
-\Cref{process} depicts the simulation process starting with matching of the data sets to receiving a valid model of the site and its jobs.
+To receive a valid model of a site and its jobs, we created the simulation process depicted in \Cref{process}.
+This process consists of five phases: Matching of data sets, grouping the data for job types, analysing the model parameters, generating a blueprint model, and lastly injecting the parameters into this model.
+Note, only a part of the data is used to calibrate the model, the other part is used for its validation.
 
@@
 	\centering
-	\includegraphics[width=0.6\linewidth]{images/process}
+	\includegraphics[width=0.55\linewidth]{images/process}
 	\caption[]{The Simulation Process}
@@
 After a job is scheduled on a resource container it acquires the specified number of job slots or waits if the node has not sufficient free slots.
-Afterwards the CPU and I/O demands of the job are processed.
+Afterwards, the CPU and I/O demands of the job are processed.
 To avoid one big I/O operation at the start of a job and to make the processing of I/O more continuously, the processing is modeled using iterations that reflect the number of events.
@@
 To simulate not explicitly modeled elements of the jobs such as the overhead created by pilot jobs we added a delay to each job.
-This delay is calibrated in order to match the measured throughput.
+This delay is calibrated in order to match the measured throughput of the test data.
 After the job finished, it releases its acquired job slots.
@@
 \begin{itemize}
-    \item We acquire Historical job monitoring information from several global CMS monitoring systems: Job reports from JobMonitoring and Workflow and Data Management framework job reports (FWJR documents) are archived on the \emph{analytix} Hadoop cluster.
+    \item We acquire historical job monitoring information from several global CMS monitoring systems: Job reports from JobMonitoring and Workflow and Data Management framework job reports (FWJR documents) are archived on the \emph{analytix} Hadoop cluster.
     The latter are being archived on the cluster as part of the WMArchive project \cite{WMArchiveWorkloadArchive2018}.
@@
 The job performance data was randomly separated into two equal sized calibration and test datasets.
+For the validation we considered the following metrics: Overall throughput of jobs, shares of the job types and the CPU efficiency of the site.
+These metrics describe how well the simulation matches the measured performance data.
 
@@
 
-Simulating one week of CMS jobs at GridKa takes about 18 minutes on a laptop.
-Thus, it would feasible to simulate several Tiers at once in a sufficient time.
-However, for simulating larger parts of the WLCG a more macroscopic model, which focuses on the interaction of the Tiers and does not model their inner structure in detail,
-is recommended.
-
-\begin{figure}
+\begin{figure}[h]
 	\vspace{-0.5ex}
 	\centering
-	\begin{tikzpicture}[scale = 1]
+	\begin{tikzpicture}[scale = 0.8]
 	\begin{axis}[
@@
 
+Simulating one week of CMS jobs at GridKa takes about 18 minutes on a laptop.
+Thus, it would feasible to simulate several Tiers at once in a sufficient time.
+However, for simulating larger parts of the WLCG a more macroscopic model, which focuses on the interaction of the Tiers and does not model their inner structure in detail,
+is recommended.
+
 \section{Related Work}
@@
 
-One possible design decision is a more sophisticated load balancing strategy that aims for a better resource utilization of the working nodes by scheduling a balanced share of CPU and IO heavy jobs to it.
+One possible design decision is a more sophisticated load balancing strategy that aims for a better resource utilization of the working nodes by scheduling a balanced share of CPU and I/O heavy jobs to it.
 To simulate this a new scheduling strategy must be implemented.

### 2018-10-02 paper/paper.tex
Kommentare AK

@@
 Furthermore, the computing resources for user jobs in HEP will be increasingly distributed and heterogeneous, resulting in a more difficult scheduling due the increasing complexity of the system.
-	A first step to meet both challenges, a more efficient utilization of the grid and coping with the rising complexity of the system, a simulation for the WLCG should be created.
-There is currently no simulation existing helping the operators of the grid to make the correct decisions while optimizing the load balancing strategy.
+	As first step to meet both challenges, a more efficient utilization of the grid and coping with the rising complexity of the system, a simulation for the WLCG should be created.
+\todo{Vorherigen Satz umbauen, Einschub im vorderen Teil schwer verständlich} There is currently no simulation existing helping the operators of the grid to make the correct decisions while optimizing the load balancing strategy.
 
-	This paper shows a proof of concept in which the computing jobs at the Tier 1 center GridKa are modeled and simulated.
+	This paper presents a proof of concept in which the computing jobs at the Tier 1 center GridKa are modeled and simulated.
 To model the computing jobs we extended the Palladio simulator with a mechanism to simulate load balancing strategies.
@@
 Cloud resources, such as HNSciCloud and AWS, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs.
-This results in a more complex scheduling, because each resource has different optimal job types and mixtures.
+This results in a more complex scheduling, because each resource has different optimal job types and mixes.
 Furthermore, this leads to a more dynamic data placement, where fewer centers host data and thus remote access is increasing.
@@
 
-To solve both challenges, a more efficient usage of the grid and coping with the rising complexity of the system, a simulation for the WLCG should be created.
+To solve both challenges, a more efficient usage of the grid and coping with the rising complexity of the system\todo{Einschub umformulieren}, a simulation for the WLCG should be created.\todo{``we aim to'' statt ``should'', auch im Abstract}
 Currently, there is no model or simulation in existence to evaluate the load balancing strategies of the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
@@
 The model and simulation are created using the Palladio simulator, see \Cref{sec:palladio} for details on Palladio.
-This leads to the research question which modifications are needed so that Palladio can be used to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio.
+This leads to the research question which modifications are needed so that Palladio can be used to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio \todo{lieber more high level: ``can we simulate?''.
+Eher: This paper presents prototype simulation model and extensions of Palladio?}.
 To realise it we created an extension for the simulator that enables us to simulate load balancing strategies and furthermore, we created an automated model parameter analysis and model creation, in order to increase the usability of the simulation.
@@
 In our proof of concept we have three data sources: The CERN WMArchive, CMS GridKa job data and GridKa node performance data.
-These data sets have to be matched because each set contains different information about an executed job, but non of them contains the complete information about it.
+These data sets have to be matched because each set contains different information about an executed job, but none of them contains the complete information about it.
 
@@
 The next step is the generation of the model of GridKa and the injection of the previously extracted performance parameters.
-Both steps can be done via the UI of Palladio and do not require manually modeling.
+Both steps can be done via the UI of Palladio and do not require manually modeling \todo{Unklar, weil UI nach manual klingt.}.
 At this point the users have a valid and calibrated model of GridKa with its resource and the workload.
@@
 
-They can then either run the simulation or make changes to the model to simulate the effect of a design change. 
+Users can then either run the simulation or make changes to the model to simulate the effect of a design change. 
 Finally, after running the simulation, they can inspect the results, such as core utilization of a node or throughput of jobs, using the integrated visualization tool or they can export the results and use a third party software to do the evaluation.
@@
 \section{Modeling Computing Jobs and Resources}
-\label{sec:model}
+\label{sec:model}\todo{Grafik?}
 This section describes how the computing jobs and working nodes are modeled using Palladio.
 The component model describes each type of CMS computing job, such as analysis or reprocessing jobs, with their resource demands.
-This demands are the CPU demand, the I/O demand, the number of used CPU cores, the required job slots, and the number of processed events.
+These demands are the CPU demand, the I/O demand, the number of used CPU cores, the required job slots, and the number of processed events.
 
-After a job is scheduled on a resource container it acquires the specified number of job slots or waits if the node has not sufficient free slots.
+After a job is scheduled on a resource container, it acquires the specified number of job slots or waits if the node has not sufficient free slots.
 Afterwards, the CPU and I/O demands of the job are processed.
@@
 \section{Model Parameter Extraction and Model Creation}
-\label{sec:param}
+\label{sec:param}\todo{Links zu den Projekten einfügen, wo langfristig archivieren? Zendodo?}
 
@@
 However, for simulating larger parts of the WLCG a more macroscopic model, which focuses on the interaction of the Tiers and does not model their inner structure in detail,
-is recommended.
+is recommended. \todo{Formulierung nicht passiv. Positiver: How to aggregate is future work?}
 

### 2018-10-04 paper/paper.tex
Implement comments

@@
 Furthermore, the computing resources for user jobs in HEP will be increasingly distributed and heterogeneous, resulting in a more difficult scheduling due the increasing complexity of the system.
-	As first step to meet both challenges, a more efficient utilization of the grid and coping with the rising complexity of the system, a simulation for the WLCG should be created.
-\todo{Vorherigen Satz umbauen, Einschub im vorderen Teil schwer verständlich} There is currently no simulation existing helping the operators of the grid to make the correct decisions while optimizing the load balancing strategy.
+	We aim to create a simulation for the WLCG helping the HEP community to solve both challenges: a more efficient utilization of the grid and coping with the rising complexity of the system.
+	There is currently no simulation existing helping the operators of the grid to make the correct decisions while optimizing the load balancing strategy.
 
@@
 
-To solve both challenges, a more efficient usage of the grid and coping with the rising complexity of the system\todo{Einschub umformulieren}, a simulation for the WLCG should be created.\todo{``we aim to'' statt ``should'', auch im Abstract}
+We aim to create a simulation for the WLCG helping the HEP community to solve both challenges: a more efficient utilization of the grid and coping with the rising complexity of the system.
 Currently, there is no model or simulation in existence to evaluate the load balancing strategies of the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
@@
 The model and simulation are created using the Palladio simulator, see \Cref{sec:palladio} for details on Palladio.
-This leads to the research question which modifications are needed so that Palladio can be used to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio \todo{lieber more high level: ``can we simulate?''.
-Eher: This paper presents prototype simulation model and extensions of Palladio?}.
+We chose Palladio to archive a scalable simulation because it is a comparably abstract simulator (\Cref{related} gives an overview on other simulators and approaches).
+This leads to the research question how we can use Palladio to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio.
 To realise it we created an extension for the simulator that enables us to simulate load balancing strategies and furthermore, we created an automated model parameter analysis and model creation, in order to increase the usability of the simulation.
@@
 The next step is the generation of the model of GridKa and the injection of the previously extracted performance parameters.
-Both steps can be done via the UI of Palladio and do not require manually modeling \todo{Unklar, weil UI nach manual klingt.}.
+Both steps do not require manually modeling.
 At this point the users have a valid and calibrated model of GridKa with its resource and the workload.
@@
 The component model describes each type of CMS computing job, such as analysis or reprocessing jobs, with their resource demands.
-These demands are the CPU demand, the I/O demand, the number of used CPU cores, the required job slots, and the number of processed events.
+These demands are the \textit{CPU demand}, the\textit{ I/O demand}, the \textit{number of used CPU cores}, the \textit{required job slot}s, and the \textit{number of processed events}.
 
@@
 Each type of working node in GridKa is modeled in the resource model.
-Each node type has a CPU and I/O processing rate, number of cores, number of job slots, and number of instances of this node type.
+Each node type has a \textit{CPU and I/O processing rate}, \textit{number of cores}, \textit{number of job slots}, and \textit{number of instances} of this node type.
 The types of computing nodes are duplicated according to their specified number of instances as a preprocessing step of the simulation, resulting in the simulation of each working node.
@@
 To make sure that there is always a high load on the system, we used a closed workload with enough jobs that the system is fully utilized.
-Each job type has a configurable share of the overall load.
+Each job type has a configurable \textit{job share} of the overall load.
 
-We extended Palladio with a \textit{load balancing strategy action}, so that we can simulate load balancing strategies.
+We extended Palladio with a \textit{load balancing strategy action}\footnote{https://github.com/PalladioSimulator/Palladio-Addon-ArchitecturalTemplates-LoadBalancing}, so that we can simulate load balancing strategies.
 This load balancing strategy action is contained in a load balancing component.
@@
 \section{Model Parameter Extraction and Model Creation}
-\label{sec:param}\todo{Links zu den Projekten einfügen, wo langfristig archivieren? Zendodo?}
+\label{sec:param}
 
@@
 For the purposes of this proof of concept, jobs were then grouped according to their type as indicated in the job reports, but the extraction framework allows to be configured for arbitrary groupings of the extracted jobs.
-Job resource requirement distributions for the different job types are extracted as distributions based on the empirical distributions in the data sets used for calibration, in the form of discretized probability density functions.
+Job resource requirement distributions for the different job types are extracted as distributions based on the empirical distributions in the data sets used for calibration, in the form of discrete probability density functions.
 
-The parameter extraction tool exports model parameter sets as JSON files.
-To simplify calibrating simulation models with these parameter sets, we created a Palladio plugin that can be used to automatically complete a blueprint of the simulation model (which defines the structure of the model as described in section \ref{sec:model}) with arbitrary parameter sets.
+The parameter extraction tool\footnote{https://github.com/mxsg/CMS-Model-Calibration} exports model parameter sets as JSON files.
+To simplify calibrating simulation models with these parameter sets, we created a Palladio plugin\footnote{https://github.com/mxsg/WLCG-Model-Construction} that can be used to automatically complete a blueprint of the simulation model (which defines the structure of the model as described in section \ref{sec:model}) with arbitrary parameter sets.
 This ensures an efficient and flexible model creation process.

### 2018-10-09 paper/paper.tex
Remove scheduling delay

@@
 While processing, the job uses the specified number of cores to process the CPU demand.
-To simulate not explicitly modeled elements of the jobs such as the overhead created by pilot jobs we added a delay to each job.
-This delay is calibrated in order to match the measured throughput of the test data.
 After the job finished, it releases its acquired job slots.
@@
 
-As mentioned in \Cref{sec:model} a delay was added to the job to simulate the effects of not explicitly modeled elements.
-This delay was calibrated so that the measured throughput matches the simulated one of the calibration dataset.
 The simulated throughput of the test dataset matches the measured one accurately with only 1\% difference.
-\Cref{shares} depicts the share of the simulated and measured job types, which also have a high accuracy.
+\Cref{shares} depicts the share of the simulated and measured job types that also shows  a high accuracy.
 

### 2018-10-09 paper/paper.tex
Add seff

@@
 \section{Modeling Computing Jobs and Resources}
-\label{sec:model}\todo{Grafik?}
+\label{sec:model}
 This section describes how the computing jobs and working nodes are modeled using Palladio.
 The component model describes each type of CMS computing job, such as analysis or reprocessing jobs, with their resource demands.
-These demands are the \textit{CPU demand}, the\textit{ I/O demand}, the \textit{number of used CPU cores}, the \textit{required job slot}s, and the \textit{number of processed events}.
+These demands are the \textit{CPU demand}, the\textit{ I/O demand}, the \textit{number of threads}, the \textit{required job slots}, and the \textit{number of demand rounds}.
 
+\Cref{seff} shows the behaviour of the job during the simulation.
 After a job is scheduled on a resource container, it acquires the specified number of job slots or waits if the node has not sufficient free slots.
 Afterwards, the CPU and I/O demands of the job are processed.
-To avoid one big I/O operation at the start of a job and to make the processing of I/O more continuously, the processing is modeled using iterations that reflect the number of events.
+To avoid one big I/O operation at the start of a job and to make the processing of I/O more continuously, the processing is modeled using iterations that reflect the number of processed events.
 Thus, for each iteration there are I/O and CPU resources needed.
 Rather than simulating each event, we simulate a fraction of the processed events, because the simulation of each event would harm the scalability of the simulation.
-While processing, the job uses the specified number of cores to process the CPU demand.
+While processing, the job uses the specified number of threads to process the CPU demand.
 After the job finished, it releases its acquired job slots.
 All of these parameters are specified as a probability function and are extracted from measured performance data as described in \Cref{sec:param}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{images/seff}
+	\caption[]{Model of Job Behaviour}
+	\label{seff}
+\end{figure}
 
@@
 
-Simulating one week of CMS jobs at GridKa takes about 18 minutes on a laptop.
-Thus, it would feasible to simulate several Tiers at once in a sufficient time.
-However, for simulating larger parts of the WLCG a more macroscopic model, which focuses on the interaction of the Tiers and does not model their inner structure in detail,
-is recommended. \todo{Formulierung nicht passiv. Positiver: How to aggregate is future work?}
+Simulating one week of CMS jobs at GridKa takes about 24 minutes on a laptop.
+Thus, it would feasible to simulate several Tiers at once in a sufficient time, however simulating larger parts of the grid is future work.
 
@@
 Furthermore, we extended Palladio to be able to simulate load balancing strategies.
-Our first results suggest that this approach could deliver accurate results regarding CPU efficiency, however it must be investigated what the reason for the data inconsistency and missing job information is.
-Regarding scalability we are confident that several computing centers could be simulated on a detailed level.
-For use cases that require the simulation of larger parts of the WLCG we suggest a more macroscopic model that focuses on the interaction of the sites and models the sites internals on a less detailed level.
+Our first results are promising and suggest that this approach can be used to simulate load balancing strategies accurately.
+Regarding scalability we are confident that several computing centers can simulated in a sufficient time, however this is part of future work.
 

### 2018-10-09 paper/paper.tex
Update template

@@
-\documentclass[a4paper]{jpconf}
+\documentclass{webofc}
+\usepackage[varg]{txfonts}
 
@@
 
-\bibliographystyle{iopart-num}
-
 \begin{document}
@@
 
-\author{Ren\'e Caspart${}^{1}$,
-	Patrick Firnkes${}^{2}$,
-	Manuel Giffels${}^{1}$,
-	Anne Koziolek${}^{2}$,
-	G\"unter Quast${}^{1}$,
-	Ralf Reussner${}^{2}$, 
-	Maximilian Stemmer-Grabow${}^{2}$}
+\author{
+	\firstname{Ren\'e} \lastname{Caspart}\inst{1} \and
+	\firstname{Patrick} \lastname{Firnkes}\inst{2} \and
+	\firstname{Manuel} \lastname{Giffels}\inst{1} \and
+	\firstname{Anne} \lastname{Koziolek}\inst{2}\fnsep\thanks{\email{koziolek@kit.edu}} \and
+	\firstname{G\"unter} \lastname{Quast}\inst{1} \and
+	\firstname{Ralf} \lastname{Reussner}\inst{2} \and
+	\firstname{Maximilian} \lastname{Stemmer-Grabow}\inst{2}
+}
 
-\address{
-	${}^{1}$Institute for Experimental Particle Physics (ETP) \\
-	${}^{2}$Institute for Program Structures and Data Organization (IPD) \\
-	at Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany}
+\institute{
+	Institute for Experimental Particle Physics (ETP) 
+	\and
+   Institute for Program Structures and Data Organization (IPD) \\
+	at Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
+}
 
-\ead{koziolek@kit.edu}
-
-\begin{abstract}
+\abstract{%
 	The amount of data to be processed by experiments in high energy physics (HEP) is tremendously increasing in the coming years.
@@
 Our results suggest that simulating larger parts of the grid is feasible and can help to optimize the utilization of the grid.
-\end{abstract}
+}%
+\maketitle
+%
 
@@
 
-\begin{figure}
+\begin{figure*}
 	\centering
@@
 	\label{resources}
-\end{figure}
+\end{figure*}
 
@@
 	\label{process}
-\end{figure}
+\end{figure*}
 
@@
 	\label{seff}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[h]
+\begin{figure*}[h]
 	\centering
@@
 	\label{shares}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[h]
+\begin{figure*}[h]
 	\vspace{-0.5ex}
@@
 	\label{utilization}
-\end{figure}
+\end{figure*}
 
@@
 The next step is the evaluation of design decisions mentioned in \Cref{outlook}.
-\section*{References}
+
 \bibliography{paper}

### 2018-10-09 paper/paper.tex
Fix typo
Rework validation section

@@
 \usepackage{pgfplots}
+
+% Use package for subfigures
+\usepackage{subcaption}
 
@@
 \label{validation}
-To validate the model and simulation we simulated CMS jobs at GridKa in April 2018.
-The job performance data was randomly separated into two equal sized calibration and test datasets.
-For the validation we considered the following metrics: Overall throughput of jobs, shares of the job types and the CPU efficiency of the site.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}{0.48\linewidth}
+        \includegraphics[width=\linewidth]{images/validation/throughputs}
+    \end{subfigure}
+    \begin{subfigure}{0.48\linewidth}
+        \includegraphics[width=\linewidth]{images/validation/utilizations}
+    \end{subfigure}
+    \caption{Measured and simulated throughputs and CPU efficiency for test data from May 2018}
+    \label{fig:validation}
+\end{figure}
+
+To validate the model and simulation we simulated CMS jobs at GridKa in May 2018.
+The job performance data was split into two equal sized calibration and test datasets by randomly sampling reports from the full data set.
+For the validation we considered the following metrics: Overall throughput of jobs, shares of the job types, and the overall CPU efficiency of the site.
 These metrics describe how well the simulation matches the measured performance data.
+Comparisons between  simulation results and measurements from the system itself for them are shown in figure \ref{fig:validation}.
+The simulated throughput matches the reference from the with a difference of about 5.6\%.
+The share of the simulated and measured job types also match accurately.
 
-The simulated throughput of the test dataset matches the measured one accurately with only 1\% difference.
-\Cref{shares} depicts the share of the simulated and measured job types that also shows  a high accuracy.
-
-\begin{figure*}[h]
-	\centering
-	\includegraphics[scale = 0.6]{images/shares}
-	\caption[]{Measured and Simulated Shares of Job Types}
-	\label{shares}
-\end{figure*}
-
-The simulated CPU efficiency is lower than the measured one, as shown in \Cref{utilization}.
-The reason for this are likely missing jobs in the job data.
-We observed that a significant amount of data is missing, the cause for this is not known yet, however this should be investigated.
-Because of that the calibrated throughput is too low, resulting in a lower utilization.
-Another reason is the missing utilization created by pilot jobs that are not explicitly modeled yet.
-
-\begin{figure*}[h]
-	\vspace{-0.5ex}
-	\centering
-	\begin{tikzpicture}[scale = 0.8]
-	\begin{axis}[
-	xlabel={Time / s},
-	ylabel={Total CPU Time / Walltime},
-	ymin=0, ymax=1,
-	xmin=0, xmax=604800,
-	legend cell align={left},
-	]
-	\addplot[color=red] coordinates {
-		(0, 0.61)			
-		(604800, 0.61)
-	};
-	\addplot[color=blue] coordinates {	
-		(14000,0.420457263335)
-		(15000,0.455205081375)
-		(16000,0.450933749429)
-		(17000,0.46853252847)
-		(18000,0.474029527868)
-		(19000,0.482875361196)
-		(20000,0.472149795914)
-		(21000,0.448813851682)
-		(22000,0.430706201872)
-		(23000,0.401599999586)
-		(24000,0.376755783861)
-		(25000,0.38749824392)
-		(26000,0.422990841095)
-		(27000,0.425760711131)
-		(28000,0.43904605249)
-		(29000,0.437695380408)
-		(30000,0.433314619731)
-		(31000,0.448449713155)
-		(32000,0.441020436611)
-		(33000,0.444429125129)
-		(34000,0.461008578928)
-		(35000,0.470284257719)
-		(36000,0.447847515795)
-		(37000,0.441187264182)
-		(38000,0.433373025865)
-		(39000,0.426778346)
-		(40000,0.440334830361)
-		(41000,0.452643935542)
-		(42000,0.452142136739)
-		(43000,0.452471374183)
-		(44000,0.445302566248)
-		(45000,0.437814297445)
-		(46000,0.445663557997)
-		(47000,0.451677101561)
-		(48000,0.442604461856)
-		(49000,0.453882226309)
-		(50000,0.444581524102)
-		(51000,0.443425493542)
-		(52000,0.424588550506)
-		(53000,0.424918434251)
-		(54000,0.438754707788)
-		(55000,0.431586617405)
-		(56000,0.432664270787)
-		(57000,0.431460959005)
-		(58000,0.43776201488)
-		(59000,0.452017801875)
-		(60000,0.445027944703)
-		(61000,0.447754439714)
-		(62000,0.441042254855)
-		(63000,0.448994039435)
-		(64000,0.447949217961)
-		(65000,0.440335544034)
-		(66000,0.432872781216)
-		(67000,0.444765956656)
-		(68000,0.445503771932)
-		(69000,0.435313479888)
-		(70000,0.440723673441)
-		(71000,0.453581869677)
-		(72000,0.455011550524)
-		(73000,0.445251024271)
-		(74000,0.461605049304)
-		(75000,0.470460077846)
-		(76000,0.467995526805)
-		(77000,0.495748712572)
-		(78000,0.49681048222)
-		(79000,0.495635118813)
-		(80000,0.476519088357)
-		(81000,0.469046301151)
-		(82000,0.459260808853)
-		(83000,0.443186601975)
-		(84000,0.453927784554)
-		(85000,0.460996975902)
-		(86000,0.439846472086)
-		(87000,0.42853879571)
-		(88000,0.425940455445)
-		(89000,0.418188095172)
-		(90000,0.441124921832)
-		(91000,0.430042869457)
-		(92000,0.454088618748)
-		(93000,0.447812967768)
-		(94000,0.472828979798)
-		(95000,0.503483458242)
-		(96000,0.49269529217)
-		(97000,0.485731894357)
-		(98000,0.500299990804)
-		(99000,0.489833797954)
-		(100000,0.486536917579)
-		(101000,0.470900103725)
-		(102000,0.473487499528)
-		(103000,0.472852199897)
-		(104000,0.465907927377)
-		(105000,0.46843843884)
-		(106000,0.469393804824)
-		(107000,0.448337382107)
-		(108000,0.453458598472)
-		(109000,0.477909863511)
-		(110000,0.484074360628)
-		(111000,0.471446789672)
-		(112000,0.483311476858)
-		(113000,0.492794346306)
-		(114000,0.494774125893)
-		(115000,0.478760296453)
-		(116000,0.463289501118)
-		(117000,0.470560118374)
-		(118000,0.46261989666)
-		(119000,0.448321324595)
-		(120000,0.4391456597)
-		(121000,0.44616088099)
-		(122000,0.443067693493)
-		(123000,0.443765615997)
-		(124000,0.45489288544)
-		(125000,0.457761907159)
-		(126000,0.459890981374)
-		(127000,0.457265336864)
-		(128000,0.450661574709)
-		(129000,0.456319671876)
-		(130000,0.455222292845)
-		(131000,0.444168136077)
-		(132000,0.455292020739)
-		(133000,0.459739659432)
-		(134000,0.474266033943)
-		(135000,0.476719078979)
-		(136000,0.454265351086)
-		(137000,0.432403127216)
-		(138000,0.426941224365)
-		(139000,0.435303571358)
-		(140000,0.445594945267)
-		(141000,0.457945319262)
-		(142000,0.467041094326)
-		(143000,0.471275880951)
-		(144000,0.448803750886)
-		(145000,0.46057752759)
-		(146000,0.450222072137)
-		(147000,0.447748079045)
-		(148000,0.463476880757)
-		(149000,0.457547341501)
-		(150000,0.464860371781)
-		(151000,0.458939834833)
-		(152000,0.449435321435)
-		(153000,0.445809287248)
-		(154000,0.452098810214)
-		(155000,0.436323981134)
-		(156000,0.435163726955)
-		(157000,0.430055068196)
-		(158000,0.432586012365)
-		(159000,0.437297849322)
-		(160000,0.424896762305)
-		(161000,0.433031154001)
-		(162000,0.42905370657)
-		(163000,0.412624449914)
-		(164000,0.412491612196)
-		(165000,0.399704545922)
-		(166000,0.407748045902)
-		(167000,0.409904586648)
-		(168000,0.412048703271)
-		(169000,0.422106417772)
-		(170000,0.431756334979)
-		(171000,0.432552532485)
-		(172000,0.461650235913)
-		(173000,0.450819280269)
-		(174000,0.455592609364)
-		(175000,0.461503577581)
-		(176000,0.463421686809)
-		(177000,0.456182986217)
-		(178000,0.465859886059)
-		(179000,0.469016662103)
-		(180000,0.457242602407)
-		(181000,0.438524914279)
-		(182000,0.419623185788)
-		(183000,0.434867384803)
-		(184000,0.436354031073)
-		(185000,0.431955984805)
-		(186000,0.426743598615)
-		(187000,0.433639371186)
-		(188000,0.447367993742)
-		(189000,0.441424080277)
-		(190000,0.460353123518)
-		(191000,0.460813912595)
-		(192000,0.455819448956)
-		(193000,0.467462336883)
-		(194000,0.475473651551)
-		(195000,0.464487515787)
-		(196000,0.45335938384)
-		(197000,0.435968082782)
-		(198000,0.459905581295)
-		(199000,0.443261087286)
-		(200000,0.438900909062)
-		(201000,0.470264321796)
-		(202000,0.451214857632)
-		(203000,0.468171752558)
-		(204000,0.459209586602)
-		(205000,0.459254298111)
-		(206000,0.47101680367)
-		(207000,0.463033170114)
-		(208000,0.4712737853)
-		(209000,0.465450336139)
-		(210000,0.455801602051)
-		(211000,0.461908888488)
-		(212000,0.458205085034)
-		(213000,0.461468054298)
-		(214000,0.453173854006)
-		(215000,0.454871547866)
-		(216000,0.444018420001)
-		(217000,0.460956695722)
-		(218000,0.465661865182)
-		(219000,0.465518305154)
-		(220000,0.4711530613)
-		(221000,0.459010429226)
-		(222000,0.464906045582)
-		(223000,0.453027764206)
-		(224000,0.455831376049)
-		(225000,0.469807813442)
-		(226000,0.456897450438)
-		(227000,0.446722893031)
-		(228000,0.444441981548)
-		(229000,0.431491369988)
-		(230000,0.448834255688)
-		(231000,0.445076241297)
-		(232000,0.435089055194)
-		(233000,0.437592817834)
-		(234000,0.437875351009)
-		(235000,0.434588735814)
-		(236000,0.44507195818)
-		(237000,0.439254948215)
-		(238000,0.444706733582)
-		(239000,0.447474042265)
-		(240000,0.446329940764)
-		(241000,0.455749636545)
-		(242000,0.419361606614)
-		(243000,0.445397279873)
-		(244000,0.452348733449)
-		(245000,0.434373580372)
-		(246000,0.43809435824)
-		(247000,0.457201403905)
-		(248000,0.458802551256)
-		(249000,0.440383453863)
-		(250000,0.445017392448)
-		(251000,0.450084408166)
-		(252000,0.441726249534)
-		(253000,0.43194266247)
-		(254000,0.445568321745)
-		(255000,0.436278081802)
-		(256000,0.439570299584)
-		(257000,0.445136501732)
-		(258000,0.45903255575)
-		(259000,0.450554481769)
-		(260000,0.446171160379)
-		(261000,0.458690240049)
-		(262000,0.470220369345)
-		(263000,0.480684091886)
-		(264000,0.456854673542)
-		(265000,0.450233371349)
-		(266000,0.457899571788)
-		(267000,0.453334820977)
-		(268000,0.460295743577)
-		(269000,0.457349902193)
-		(270000,0.462039377481)
-		(271000,0.481382348042)
-		(272000,0.491382702182)
-		(273000,0.484220311688)
-		(274000,0.491652039182)
-		(275000,0.486379984979)
-		(276000,0.494978853484)
-		(277000,0.490824451822)
-		(278000,0.461792687666)
-		(279000,0.469822396762)
-		(280000,0.48576001703)
-		(281000,0.481488945743)
-		(282000,0.480743747455)
-		(283000,0.486090522042)
-		(284000,0.487356854396)
-		(285000,0.465745257924)
-		(286000,0.472330225888)
-		(287000,0.492995112907)
-		(288000,0.489510571313)
-		(289000,0.493418264292)
-		(290000,0.489755865611)
-		(291000,0.489256145131)
-		(292000,0.483996042192)
-		(293000,0.471077668813)
-		(294000,0.471398859559)
-		(295000,0.461127782716)
-		(296000,0.458754727805)
-		(297000,0.451312728894)
-		(298000,0.457666739683)
-		(299000,0.463072024695)
-		(300000,0.452352488238)
-		(301000,0.432056039316)
-		(302000,0.442534543126)
-		(303000,0.464020942755)
-		(304000,0.465788968846)
-		(305000,0.475194751884)
-		(306000,0.460429089784)
-		(307000,0.440203075328)
-		(308000,0.440392507843)
-		(309000,0.469031285375)
-		(310000,0.457737564909)
-		(311000,0.450884070739)
-		(312000,0.467259163074)
-		(313000,0.467018061082)
-		(314000,0.466325864949)
-		(315000,0.4515369024)
-		(316000,0.451628960618)
-		(317000,0.449715736896)
-		(318000,0.453599972995)
-		(319000,0.46598650055)
-		(320000,0.458071641171)
-		(321000,0.459795875349)
-		(322000,0.463776572273)
-		(323000,0.441893234874)
-		(324000,0.450990407649)
-		(325000,0.469838695448)
-		(326000,0.454971976955)
-		(327000,0.458423340158)
-		(328000,0.455021879641)
-		(329000,0.474962772272)
-		(330000,0.468921434665)
-		(331000,0.463586286578)
-		(332000,0.472336338416)
-		(333000,0.480203877337)
-		(334000,0.499976146642)
-		(335000,0.499528976694)
-		(336000,0.486582556042)
-		(337000,0.486946662392)
-		(338000,0.472474569197)
-		(339000,0.47914245784)
-		(340000,0.475286986461)
-		(341000,0.458659120511)
-		(342000,0.463977117388)
-		(343000,0.48570346922)
-		(344000,0.47347942112)
-		(345000,0.450392499597)
-		(346000,0.445409073982)
-		(347000,0.454343590968)
-		(348000,0.446283787024)
-		(349000,0.441991195931)
-		(350000,0.457080478112)
-		(351000,0.475143781352)
-		(352000,0.453825073455)
-		(353000,0.466696828065)
-		(354000,0.461398654993)
-		(355000,0.476432737168)
-		(356000,0.484493658032)
-		(357000,0.478381386037)
-		(358000,0.48791767455)
-		(359000,0.471135835878)
-		(360000,0.4510006391)
-		(361000,0.459254358243)
-		(362000,0.45940296355)
-		(363000,0.455276346902)
-		(364000,0.460372976975)
-		(365000,0.46432780643)
-		(366000,0.468703252372)
-		(367000,0.460690627792)
-		(368000,0.456596843643)
-		(369000,0.453188747838)
-		(370000,0.444026572421)
-		(371000,0.469540885626)
-		(372000,0.477111174668)
-		(373000,0.453778831255)
-		(374000,0.451432024952)
-		(375000,0.467745860758)
-		(376000,0.488699820134)
-		(377000,0.47828379946)
-		(378000,0.470780021427)
-		(379000,0.460858540737)
-		(380000,0.468116574189)
-		(381000,0.476739105023)
-		(382000,0.465550165705)
-		(383000,0.466704565487)
-		(384000,0.460951873868)
-		(385000,0.474085333555)
-		(386000,0.483774319139)
-		(387000,0.450257163363)
-		(388000,0.432558156082)
-		(389000,0.441761962984)
-		(390000,0.471357238783)
-		(391000,0.476402662876)
-		(392000,0.478773054016)
-		(393000,0.479343982261)
-		(394000,0.476955005539)
-		(395000,0.464686525062)
-		(396000,0.465867162002)
-		(397000,0.473713210956)
-		(398000,0.455977533182)
-		(399000,0.453556194091)
-		(400000,0.468032168961)
-		(401000,0.473372878696)
-		(402000,0.460103363611)
-		(403000,0.450437880682)
-		(404000,0.444189180728)
-		(405000,0.440631553941)
-		(406000,0.440814465656)
-		(407000,0.439132208745)
-		(408000,0.443680589194)
-		(409000,0.439364305134)
-		(410000,0.445754620988)
-		(411000,0.459675216489)
-		(412000,0.434411774646)
-		(413000,0.460859998054)
-		(414000,0.470068659819)
-		(415000,0.46398628613)
-		(416000,0.46092602178)
-		(417000,0.470952398074)
-		(418000,0.464918768054)
-		(419000,0.457250642935)
-		(420000,0.452696999147)
-		(421000,0.443597876602)
-		(422000,0.442688896448)
-		(423000,0.446975959689)
-		(424000,0.459962628411)
-		(425000,0.458639133149)
-		(426000,0.45869214177)
-		(427000,0.459339918847)
-		(428000,0.469954533509)
-		(429000,0.466801073197)
-		(430000,0.479585876798)
-		(431000,0.473889297072)
-		(432000,0.468945113898)
-		(433000,0.46819782837)
-		(434000,0.460713604747)
-		(435000,0.467085533437)
-		(436000,0.464877942532)
-		(437000,0.441101434716)
-		(438000,0.435732064901)
-		(439000,0.448217580287)
-		(440000,0.447917450562)
-		(441000,0.443407621096)
-		(442000,0.443456810485)
-		(443000,0.438987312155)
-		(444000,0.467286398042)
-		(445000,0.469189496526)
-		(446000,0.467728960254)
-		(447000,0.469175237549)
-		(448000,0.469335300568)
-		(449000,0.463820736067)
-		(450000,0.458428492443)
-		(451000,0.448441949348)
-		(452000,0.453900123964)
-		(453000,0.459953017199)
-		(454000,0.448381040252)
-		(455000,0.465700695725)
-		(456000,0.451531261356)
-		(457000,0.456563899123)
-		(458000,0.45720539038)
-		(459000,0.451801077935)
-		(460000,0.464657248779)
-		(461000,0.459373104135)
-		(462000,0.469958682024)
-		(463000,0.467903777714)
-		(464000,0.445428785915)
-		(465000,0.442499436959)
-		(466000,0.447134373326)
-		(467000,0.438282409091)
-		(468000,0.43839540036)
-		(469000,0.419945925767)
-		(470000,0.418452547982)
-		(471000,0.438307200322)
-		(472000,0.45350446074)
-		(473000,0.468786686762)
-		(474000,0.456254084131)
-		(475000,0.453297591133)
-		(476000,0.469078411526)
-		(477000,0.457354866601)
-		(478000,0.476430954364)
-		(479000,0.475029005373)
-		(480000,0.479985350112)
-		(481000,0.475381159607)
-		(482000,0.461543267652)
-		(483000,0.451592900155)
-		(484000,0.45260612709)
-		(485000,0.458008617252)
-		(486000,0.444827616864)
-		(487000,0.447400533469)
-		(488000,0.452399703406)
-		(489000,0.451208289224)
-		(490000,0.469255926366)
-		(491000,0.469056554934)
-		(492000,0.448427595423)
-		(493000,0.460791545132)
-		(494000,0.465807144119)
-		(495000,0.459466948041)
-		(496000,0.457398936667)
-		(497000,0.454270854364)
-		(498000,0.44030861198)
-		(499000,0.441478571167)
-		(500000,0.448416540804)
-		(501000,0.447708055398)
-		(502000,0.44394661343)
-		(503000,0.437604010563)
-		(504000,0.43279917301)
-		(505000,0.448202463548)
-		(506000,0.450335081656)
-		(507000,0.456367109826)
-		(508000,0.476797822515)
-		(509000,0.464004036098)
-		(510000,0.447462462599)
-		(511000,0.460128351384)
-		(512000,0.456610531662)
-		(513000,0.461027700674)
-		(514000,0.452647871516)
-		(515000,0.472383445652)
-		(516000,0.480048107691)
-		(517000,0.449495011852)
-		(518000,0.424921363583)
-		(519000,0.430957570675)
-		(520000,0.42946792577)
-		(521000,0.435679220854)
-		(522000,0.454543429655)
-		(523000,0.449873483822)
-		(524000,0.441368528175)
-		(525000,0.434726970329)
-		(526000,0.446558845708)
-		(527000,0.436523566317)
-		(528000,0.436567752145)
-		(529000,0.440939357596)
-		(530000,0.442182493949)
-		(531000,0.446092107838)
-		(532000,0.451187916749)
-		(533000,0.439896893747)
-		(534000,0.449011137767)
-		(535000,0.444120664286)
-		(536000,0.426232253223)
-		(537000,0.438918286378)
-		(538000,0.44181502269)
-		(539000,0.432995321972)
-		(540000,0.434879809332)
-		(541000,0.427145288337)
-		(542000,0.427872289185)
-		(543000,0.417701942996)
-		(544000,0.419515873503)
-		(545000,0.439790579048)
-		(546000,0.43762292406)
-		(547000,0.443042538052)
-		(548000,0.455727191114)
-		(549000,0.447614026634)
-		(550000,0.456624751487)
-		(551000,0.452793939028)
-		(552000,0.440726166681)
-		(553000,0.449038613212)
-		(554000,0.468536381832)
-		(555000,0.444002288289)
-		(556000,0.452092079245)
-		(557000,0.456883014276)
-		(558000,0.453112451125)
-		(559000,0.441037791264)
-		(560000,0.442347996342)
-		(561000,0.436264841204)
-		(562000,0.438959132682)
-		(563000,0.4451724287)
-		(564000,0.458215756554)
-		(565000,0.448690251696)
-		(566000,0.453197454907)
-		(567000,0.453794387302)
-		(568000,0.444846550094)
-		(569000,0.449982902679)
-		(570000,0.447267180536)
-		(571000,0.445828406211)
-		(572000,0.446909040851)
-		(573000,0.453980157655)
-		(574000,0.457292498947)
-		(575000,0.448977443253)
-		(576000,0.446078471261)
-		(577000,0.434934699022)
-		(578000,0.45949539644)
-		(579000,0.477069726029)
-		(580000,0.478983629734)
-		(581000,0.459161976806)
-		(582000,0.470964357654)
-		(583000,0.466734817389)
-		(584000,0.451817556376)
-		(585000,0.438500345769)
-		(586000,0.435540269671)
-		(587000,0.436722886274)
-		(588000,0.452696266979)
-		(589000,0.444922652517)
-		(590000,0.414546052368)
-		(591000,0.40666087591)
-		(592000,0.431870620823)
-		(593000,0.456065081956)
-		(594000,0.45959389509)
-		(595000,0.445709712809)
-		(596000,0.455310403564)
-		(597000,0.474126321375)
-		(598000,0.466945015295)
-		(599000,0.442916390259)
-		(600000,0.463327317157)
-		(601000,0.463538228748)
-		(602000,0.484414072165)
-		(603000,0.470969958228)
-		(604000,0.357975119833)
-
-	};
-	\addlegendentry{Measured (average)};
-	\addlegendentry{Simulated};
-	\end{axis}
-	\end{tikzpicture}
-	\caption[]{Measured and Simulated CPU Efficiency}
-	\label{utilization}
-\end{figure*}
+The CPU utilization in the simulation and a comparison with the measured CPU efficiency is also shown in figure \ref{fig:validation}.
+Here, the ramp-up of the simulation is also visible.
+The overall average utilization in the simulation deviates from the reference by about 3\%, showing that even the current simplified model is capable of accurately reproducing the metrics introduced above.
+Runtimes of the simulated jobs were also measured, the average runtime of simulated jobs deviates by approximately 8\% from the measured runtimes in the test data set.
 
 Simulating one week of CMS jobs at GridKa takes about 24 minutes on a laptop.
-Thus, it would feasible to simulate several Tiers at once in a sufficient time, however simulating larger parts of the grid is future work.
+Thus, it would be feasible to simulate several sites at once in a reasonable time, however simulating larger parts of the grid is future work.
 
@@
 Our first results are promising and suggest that this approach can be used to simulate load balancing strategies accurately.
-Regarding scalability we are confident that several computing centers can simulated in a sufficient time, however this is part of future work.
+Regarding scalability we are confident that several computing centers can be simulated in a sufficient time, however this is part of future work.
 

### 2018-10-09 paper/paper.tex
Minor corrections

@@
 To cope with this increasing load, most efficient usage of the resources is mandatory.
-Furthermore, the computing resources for user jobs in HEP will be increasingly distributed and heterogeneous, resulting in a more difficult scheduling due the increasing complexity of the system.
+Furthermore, the computing resources for user jobs in HEP will be increasingly distributed and heterogeneous, resulting in more difficult scheduling due the increasing complexity of the system.
 	We aim to create a simulation for the WLCG helping the HEP community to solve both challenges: a more efficient utilization of the grid and coping with the rising complexity of the system.
@@
 
-The amount of data processed by experiments in high energy physics (HEP) will drastically increase in the coming years due the upgrade of the LHC to the High-Luminosity Large Hadron Collider (HL-LHC), creating major challenges for the community of HEP \cite{community}.
+The amount of data processed by experiments in high energy physics (HEP) will drastically increase in the coming years due the upgrade of the LHC to the High-Luminosity Large Hadron Collider (HL-LHC), creating major challenges for the HEP community \cite{community}.
 \Cref{resources} shows the estimated required and provided CPU resources for the ATLAS experiment in the next ten years.
@@
 Afterwards, the CPU and I/O demands of the job are processed.
-To avoid one big I/O operation at the start of a job and to make the processing of I/O more continuously, the processing is modeled using iterations that reflect the number of processed events.
+To avoid one big I/O operation at the start of a job and to make the simulated processing of I/O more continuous, it is modeled using iterations that reflect the number of processed events.
 Thus, for each iteration there are I/O and CPU resources needed.
@@
     \end{subfigure}
-    \caption{Measured and simulated throughputs and CPU efficiency for test data from May 2018}
+    \caption{Measured and simulated throughputs and CPU efficiency, job data from May 2018}
     \label{fig:validation}
@@
 Comparisons between  simulation results and measurements from the system itself for them are shown in figure \ref{fig:validation}.
-The simulated throughput matches the reference from the with a difference of about 5.6\%.
+The simulated throughput matches the reference with a difference of about 5.6\%.
 The share of the simulated and measured job types also match accurately.
@@
 The CPU utilization in the simulation and a comparison with the measured CPU efficiency is also shown in figure \ref{fig:validation}.
-Here, the ramp-up of the simulation is also visible.
-The overall average utilization in the simulation deviates from the reference by about 3\%, showing that even the current simplified model is capable of accurately reproducing the metrics introduced above.
+The initial ramp-up of the simulation is also visible.
+The overall average utilization in the simulation deviates from the reference by about 3\%, showing that even the current simplified model is capable of reproducing the metrics introduced above.
 Runtimes of the simulated jobs were also measured, the average runtime of simulated jobs deviates by approximately 8\% from the measured runtimes in the test data set.

### 2018-10-09 paper/paper.tex
Add missing word

@@
 To cope with this increasing load, most efficient usage of the resources is mandatory.
-Furthermore, the computing resources for user jobs in HEP will be increasingly distributed and heterogeneous, resulting in more difficult scheduling due the increasing complexity of the system.
+Furthermore, the computing resources for user jobs in HEP will be increasingly distributed and heterogeneous, resulting in more difficult scheduling due to the increasing complexity of the system.
 	We aim to create a simulation for the WLCG helping the HEP community to solve both challenges: a more efficient utilization of the grid and coping with the rising complexity of the system.

### 2018-10-10 paper/paper.tex
Change layout to fit in 8pages

@@
 \usepackage{subcaption}
+
+% layout changes
+\setlength\belowcaptionskip{-15pt}
+\renewcommand*{\bibfont}{\footnotesize}
 
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{images/resources}
+	\includegraphics[width=0.5\linewidth]{images/resources}
 	\caption[]{Estimated CPU resources of ATLAS \cite{community}}
@@
 This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG.
-Furthermore, it can be used as a foundation for models and simulations of larger parts of the WLCG.
+Also, it can be used as a foundation for models and simulations of larger parts of the WLCG.
 
@@
 	\centering
-	\includegraphics[width=0.55\linewidth]{images/process}
+	\includegraphics[width=0.45\linewidth]{images/process}
 	\caption[]{The Simulation Process}
@@
 	\centering
-	\includegraphics[width=1\linewidth]{images/seff}
-	\caption[]{Model of Job Behaviour}
+	\includegraphics[width=0.8\linewidth]{images/seff}
+	\caption[]{Model of Job Behavior}
 	\label{seff}
@@
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
+	\vspace{-10px}
 	\begin{subfigure}{0.48\linewidth}
@@
 	\end{subfigure}
+	\vspace{-10px}
 	\caption{Measured and simulated throughputs and CPU efficiency, job data from May 2018}
@@
 \end{figure}
-
-To validate the model and simulation we simulated CMS jobs at GridKa in May 2018.
+\vspace{10px}
+\noindent To validate the model and simulation we simulated CMS jobs at GridKa in May 2018.
 The job performance data was split into two equal sized calibration and test datasets by randomly sampling reports from the full data set.
@@
 These metrics describe how well the simulation matches the measured performance data.
-Comparisons between  simulation results and measurements from the system itself for them are shown in figure \ref{fig:validation}.
+Comparisons between  simulation results and measurements from the system itself for them are shown in figure \Cref{fig:validation}.
 The simulated throughput matches the reference with a difference of about 5.6\%.
@@
 
-The CPU utilization in the simulation and a comparison with the measured CPU efficiency is also shown in figure \ref{fig:validation}.
+The CPU utilization in the simulation and a comparison with the measured CPU efficiency is also shown in figure \Cref{fig:validation}.
 The initial ramp-up of the simulation is also visible.
@@
 Simulating one week of CMS jobs at GridKa takes about 24 minutes on a laptop.
-Thus, it would be feasible to simulate several sites at once in a reasonable time, however simulating larger parts of the grid is future work.
+Thus, it is feasible to simulate several sites at once in a reasonable time, however simulating larger parts of the grid is future work.
 

### 2018-10-17 paper/paper.tex
Increase text front of seff
Small corrections

@@
 \author{
-	\firstname{Ren\'e} \lastname{Caspart}\inst{1} \and
-	\firstname{Patrick} \lastname{Firnkes}\inst{2} \and
-	\firstname{Manuel} \lastname{Giffels}\inst{1} \and
+	\firstname{Ren\'e} \lastname{Caspart}\inst{1}\fnsep\thanks{\email{rene.caspart@cern.ch}} \and
+	\firstname{Patrick} \lastname{Firnkes}\inst{2}\fnsep\thanks{\email{patrick.firnkes@student.kit.edu}} \and
+	\firstname{Manuel} \lastname{Giffels}\inst{1}\fnsep\thanks{\email{manuel.giffels@cern.ch}} \and
 	\firstname{Anne} \lastname{Koziolek}\inst{2}\fnsep\thanks{\email{koziolek@kit.edu}} \and
-	\firstname{G\"unter} \lastname{Quast}\inst{1} \and
-	\firstname{Ralf} \lastname{Reussner}\inst{2} \and
-	\firstname{Maximilian} \lastname{Stemmer-Grabow}\inst{2}
+	\firstname{G\"unter} \lastname{Quast}\inst{1}\fnsep\thanks{\email{g.quast@kit.edu}} \and
+	\firstname{Ralf} \lastname{Reussner}\inst{2}\fnsep\thanks{\email{reussner@kit.edu}} \and
+	\firstname{Maximilian} \lastname{Stemmer-Grabow}\fnsep\thanks{\email{mail@mxsg.de}}\inst{2}
 }
@@
 The model and simulation are created using the Palladio simulator, see \Cref{sec:palladio} for details on Palladio.
-We chose Palladio to archive a scalable simulation because it is a comparably abstract simulator (\Cref{related} gives an overview on other simulators and approaches).
+We chose Palladio to achieve a scalable simulation because it is a comparably abstract simulator (\Cref{related} gives an overview on other simulators and approaches).
 This leads to the research question how we can use Palladio to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio.
 To realise it we created an extension for the simulator that enables us to simulate load balancing strategies and furthermore, we created an automated model parameter analysis and model creation, in order to increase the usability of the simulation.
-This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG.
+This model makes it possible to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG.
 Also, it can be used as a foundation for models and simulations of larger parts of the WLCG.
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{images/seff}
+	\includegraphics[width=0.82\linewidth]{images/seff}
 	\caption[]{Model of Job Behavior}
@@
 These metrics describe how well the simulation matches the measured performance data.
-Comparisons between  simulation results and measurements from the system itself for them are shown in figure \Cref{fig:validation}.
-The simulated throughput matches the reference with a difference of about 5.6\%.
-The share of the simulated and measured job types also match accurately.
+Comparisons between simulation results and measurements from the system are shown in \Cref{fig:validation}.
+One result is that simulated throughput matches the reference with a difference of about 5.6\%.
+Furthermore, the shares of the simulated and measured job types also match accurately.
 
-The CPU utilization in the simulation and a comparison with the measured CPU efficiency is also shown in figure \Cref{fig:validation}.
-The initial ramp-up of the simulation is also visible.
+The CPU utilization in the simulation and a comparison with the measured CPU efficiency are shown in \Cref{fig:validation}, which also makes the initial ramp-up of the simulation visible.
 The overall average utilization in the simulation deviates from the reference by about 3\%, showing that even the current simplified model is capable of reproducing the metrics introduced above.
-Runtimes of the simulated jobs were also measured, the average runtime of simulated jobs deviates by approximately 8\% from the measured runtimes in the test data set.
+Additionally, the runtimes of the simulated jobs were measured: The average runtime of simulated jobs deviates by approximately 8\% from the measured runtimes in the test data set.
 
@@
 Secondly, the jobs must be parametrized with the possibility of remote access.
-Then the data cache should be modeled.
+Afterwards, the data cache with its hit rate should be modeled.
 Finally, the performance improvements of the data cache by transferring data from the cache and not from sites can be simulated.

### 2018-10-17 paper/paper.tex
Small layout changes
Use double colum for footnotes and replace github url with doi url.

@@
 \usepackage{pgfplots}
-
+\newcommand{\footurl}[1]{\footnote{\url{#1}}}
 % Use package for subfigures
@@
 % layout changes
+\usepackage{dblfnote}
+\setlength\abovecaptionskip{5pt}
 \setlength\belowcaptionskip{-15pt}
@@
 	\firstname{Ralf} \lastname{Reussner}\inst{2}\fnsep\thanks{\email{reussner@kit.edu}} \and
-	\firstname{Maximilian} \lastname{Stemmer-Grabow}\fnsep\thanks{\email{mail@mxsg.de}}\inst{2}
+	\firstname{Maximilian} \lastname{Stemmer-Grabow}\inst{2}\fnsep\thanks{\email{mail@mxsg.de}}
 }
@@
 
-We extended Palladio with a \textit{load balancing strategy action}\footnote{https://github.com/PalladioSimulator/Palladio-Addon-ArchitecturalTemplates-LoadBalancing}, so that we can simulate load balancing strategies.
+We extended Palladio with a \textit{load balancing strategy action}\footurl{https://doi.org/10.5281/zenodo.1464569}, so that we can simulate load balancing strategies.
 This load balancing strategy action is contained in a load balancing component.
@@
 
-The parameter extraction tool\footnote{https://github.com/mxsg/CMS-Model-Calibration} exports model parameter sets as JSON files.
-To simplify calibrating simulation models with these parameter sets, we created a Palladio plugin\footnote{https://github.com/mxsg/WLCG-Model-Construction} that can be used to automatically complete a blueprint of the simulation model (which defines the structure of the model as described in section \ref{sec:model}) with arbitrary parameter sets.
+The parameter extraction tool\footurl{https://doi.org/10.5281/zenodo.1464569} exports model parameter sets as JSON files.
+To simplify calibrating simulation models with these parameter sets, we created a Palladio plugin\footurl{https://doi.org/10.5281/zenodo.1464569} that can be used to automatically complete a blueprint of the simulation model (which defines the structure of the model as described in section \ref{sec:model}) with arbitrary parameter sets.
 This ensures an efficient and flexible model creation process.

### 2018-10-18 paper/paper.tex
Fix caption style
Respect page limit of 8 pages
Update Zenodo DOI links

@@
 \newcommand{\footurl}[1]{\footnote{\url{#1}}}
+
 % Use package for subfigures
 \usepackage{subcaption}
+\captionsetup[figure]{font={normalfont, small}, labelfont={bf}, labelsep=period, justification=raggedright,singlelinecheck=false}
 
@@
 \usepackage{dblfnote}
+
 \setlength\abovecaptionskip{5pt}
@@
 	Institute for Experimental Particle Physics (ETP) 
-	\and
-   Institute for Program Structures and Data Organization (IPD) \\
+	\and Institute for Program Structures and Data Organization (IPD) \\
 	at Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
@@
 performance) using several models of a system \cite{BECKER20093}.
-
-These models are the component model, the assembly model, the resource model, the allocation model, and the usage model.
+These models are:
 \begin{itemize}
@@
 	\centering
-	\includegraphics[width=0.45\linewidth]{images/process}
+	\includegraphics[width=0.39\linewidth]{images/process}
 	\caption[]{The Simulation Process}
@@
 
-The parameter extraction tool\footurl{https://doi.org/10.5281/zenodo.1464569} exports model parameter sets as JSON files.
-To simplify calibrating simulation models with these parameter sets, we created a Palladio plugin\footurl{https://doi.org/10.5281/zenodo.1464569} that can be used to automatically complete a blueprint of the simulation model (which defines the structure of the model as described in section \ref{sec:model}) with arbitrary parameter sets.
+The parameter extraction tool\footurl{https://doi.org/10.5281/zenodo.1466003} exports model parameter sets as JSON files.
+To simplify calibrating simulation models with these parameter sets, we created a Palladio plugin\footurl{https://doi.org/10.5281/zenodo.1466006} that can be used to automatically complete a blueprint of the simulation model (which defines the structure of the model as described in section \ref{sec:model}) with arbitrary parameter sets.
 This ensures an efficient and flexible model creation process.

### 2018-10-22 paper/paper.tex
Some linguistic improvements

@@
 \abstract{%
-	The amount of data to be processed by experiments in high energy physics (HEP) is tremendously increasing in the coming years.
+	The amount of data to be processed by experiments in high energy physics (HEP) will increase tremendously in the coming years.
 To cope with this increasing load, most efficient usage of the resources is mandatory.
@@
 	We aim to create a simulation for the WLCG helping the HEP community to solve both challenges: a more efficient utilization of the grid and coping with the rising complexity of the system.
-	There is currently no simulation existing helping the operators of the grid to make the correct decisions while optimizing the load balancing strategy.
+	There is currently no simulation in existence which helps the operators of the grid to make the correct decisions while optimizing the load balancing strategy.
 
@@
 
-	Finally, the simulation results are validated using real word performance data.
+	Finally, the simulation results are validated using real-word performance data.
 Our results suggest that simulating larger parts of the grid is feasible and can help to optimize the utilization of the grid.
@@
 \Cref{related} describes different attempts to simulate the WLCG or parts of it.
-However, they showed scalability issues or the simulations do not have the necessary detail level to simulate effects on working nodes that is necessary to evaluate load balancing strategies.
+However, they either showed scalability issues or the simulations do not have the necessary detail level to simulate effects on working nodes that is necessary to evaluate load balancing strategies.
 
-Optimizing load balancing strategies without the help of simulations requires stable load, in order to preserve the comparability of the different strategies.
+Optimizing load balancing strategies without the help of simulations requires stable load in order to preserve the comparability of the different strategies.
 This is neither possible to achieve for the WLCG nor for a single center like GridKa, because the nature of the load is highly dynamical.
 Another option would be to use an isolated test system for the evaluation.
-However, the system under inspection is too large and thus a similar test systems would not be financially affordable.
-A smaller test system could not deliver correct results because too much of the complexity of the system would be taken away.
+However, the system under inspection is too large and thus a similar test system would not be financially affordable.
+A smaller test system could not deliver the correct results because too much of the complexity of the system would be taken away.
 These reasons lead to the conclusion that a model and simulation of the performance of computing jobs is required to evaluate different load balancing strategies.
@@
 We chose Palladio to achieve a scalable simulation because it is a comparably abstract simulator (\Cref{related} gives an overview on other simulators and approaches).
-This leads to the research question how we can use Palladio to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio.
-To realise it we created an extension for the simulator that enables us to simulate load balancing strategies and furthermore, we created an automated model parameter analysis and model creation, in order to increase the usability of the simulation.
+This leads to the research question of how we can use Palladio to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio.
+To answer this question, we created an extension for the simulator that enables us to simulate load balancing strategies.
+Furthermore, we created an automated model parameter analysis and model creation in order to increase the usability of the simulation.
 This model makes it possible to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG.
@@
 \Cref{sec:param} explains how we determine the model parameters and the automated model creation works.
-The validation is presented in \Cref{validation} and \Cref{related} puts our work in context with the related work.
-In \Cref{outlook} possible design decisions and how they could be modeled and simulated are presented.
+The validation is presented in \Cref{validation} and \Cref{related} puts our work into context with the related work.
+In \Cref{outlook}, possible design decisions and how they could be modeled and simulated, are presented.
 Finally, \Cref{conlusion} gives a conclusion on this paper.
@@
 To receive a valid model of a site and its jobs, we created the simulation process depicted in \Cref{process}.
-This process consists of five phases: Matching of data sets, grouping the data for job types, analysing the model parameters, generating a blueprint model, and lastly injecting the parameters into this model.
+This process consists of five phases: Matching of data sets, grouping the data for job types, analysing the model parameters, generating a blueprint model, and finally injecting the parameters into this model.
 Note, only a part of the data is used to calibrate the model, the other part is used for its validation.
@@
 In our proof of concept we have three data sources: The CERN WMArchive, CMS GridKa job data and GridKa node performance data.
-These data sets have to be matched because each set contains different information about an executed job, but none of them contains the complete information about it.
+These data sets have to be matched because each set contains different information about an executed job, but none contain the complete information about it.
 
@@
 The next step is the generation of the model of GridKa and the injection of the previously extracted performance parameters.
-Both steps do not require manually modeling.
+Neither step requires manually modeling.
 At this point the users have a valid and calibrated model of GridKa with its resource and the workload.
@@
 \Cref{seff} shows the behaviour of the job during the simulation.
-After a job is scheduled on a resource container, it acquires the specified number of job slots or waits if the node has not sufficient free slots.
+After a job is scheduled on a resource container, it acquires the specified number of job slots or waits if the node does not have sufficient free slots.
 Afterwards, the CPU and I/O demands of the job are processed.
@@
 Rather than simulating each event, we simulate a fraction of the processed events, because the simulation of each event would harm the scalability of the simulation.
-While processing, the job uses the specified number of threads to process the CPU demand.
-After the job finished, it releases its acquired job slots.
+Whilst processing, the job uses the specified number of threads to process the CPU demand.
+Once the job is finished, it releases its acquired job slots.
 All of these parameters are specified as a probability function and are extracted from measured performance data as described in \Cref{sec:param}.
@@
 
-At GridKa every type of computing job can run on every node, thus each job type is allocated on each computing node in the allocation model.
+At GridKa every type of computing job can run on every node, thus each job type is allocated to each computing node in the allocation model.
 
@@
 In the assembly model all job components are connected with the load balancing component, which is the entry point of the system.
-This action decides on which node a job should be scheduled based on the currently free job slots of each working node. The load balancing strategies can be extended to evaluate different strategies.
+This action decides to which node a job should be scheduled, based on the currently free job slots of each working node. The load balancing strategies can be extended to evaluate different strategies.
 
@@
     \item We acquire historical job monitoring information from several global CMS monitoring systems: Job reports from JobMonitoring and Workflow and Data Management framework job reports (FWJR documents) are archived on the \emph{analytix} Hadoop cluster.
-    The latter are being archived on the cluster as part of the WMArchive project \cite{WMArchiveWorkloadArchive2018}.
+    The latter are archived on the cluster as part of the WMArchive project \cite{WMArchiveWorkloadArchive2018}.
     We collect job information using the \emph{CMSSpark} framework by Kuznetsov et al. \cite{cmsspark}.
@@
 One result is that simulated throughput matches the reference with a difference of about 5.6\%.
-Furthermore, the shares of the simulated and measured job types also match accurately.
+Furthermore, the shares of the simulated and measured job types also match closely.
 
@@
 Simulating one week of CMS jobs at GridKa takes about 24 minutes on a laptop.
-Thus, it is feasible to simulate several sites at once in a reasonable time, however simulating larger parts of the grid is future work.
+Thus, it is feasible to simulate several sites at once in a reasonable time, however simulating larger parts of the grid is left as future work.
 
@@
 
-There are a variety of grid and cloud simulators existing, of which SimGrid and CloudSim are two popular ones. 
+There are a variety of grid and cloud simulators in existence today, of which SimGrid and CloudSim are two popular examples. 
 However, both of them are implemented as libraries and not as standalone simulators.
@@
 The support of graphical editors is a big advantage of Palladio and increases its usability and efficiency.
-Another advantage of Palladio is the separation of the system into different models, which enables one to change on part of the system without modifying others.
-Furthermore, both do lack the support of an integrated visualization tool to display the results.
+Another advantage of Palladio is the separation of the system into different models, which enables one to change one part of the system without modifying others.
+Furthermore, both lack the support of an integrated visualization tool to display the results.
 
@@
 The Prediction Toolkit is built on Palladio and SimuLizar.
-A first case studies in the context of chemical computing showed that the Prediction Toolkit has a high accuracy, but it was not tested on large systems yet \cite{rapidtesting}.
+A first case study in the context of chemical computing showed that the Prediction Toolkit has high accuracy, but has not been tested on large systems yet \cite{rapidtesting}.
 However, due the required Runtime Toolkit CACTOS was not applicable in our use case.
@@
 
-One possible design decision is a more sophisticated load balancing strategy that aims for a better resource utilization of the working nodes by scheduling a balanced share of CPU and I/O heavy jobs to it.
-To simulate this a new scheduling strategy must be implemented.
+One possible design decision is a more sophisticated load balancing strategy that aims for a better resource utilization of the working nodes by scheduling to it a balanced share of CPU and I/O heavy jobs.
+To simulate this, a new scheduling strategy must be implemented.
 
-Another design decision would be to experiment with different sizes of pilot jobs, assuming that smaller pilot jobs result in a better utilization due less wasted CPU capacity in the draining phase of a pilot.
+Another design decision would be to experiment with different sizes of pilot jobs, assuming that smaller pilot jobs result in a better utilization due to less wasted CPU capacity in the draining phase of a pilot.
 For this use case the model must be extended with the explicit modeling of pilots jobs.
@@
 The model must be extended with the cloud resources and the required network load to access the files from a site.
-Then a good load balancing strategy for the  integration of cloud resources could be evaluated, like only scheduling CPU heavy jobs to cloud resources and keeping I/O heavy jobs near the data.
+Then a good load balancing strategy for the  integration of cloud resources could be evaluated, such as only scheduling CPU heavy jobs to cloud resources and keeping I/O heavy jobs near the data.
 
@@
 Secondly, the jobs must be parametrized with the possibility of remote access.
-Afterwards, the data cache with its hit rate should be modeled.
-Finally, the performance improvements of the data cache by transferring data from the cache and not from sites can be simulated.
+Thirdly, the data cache with its hit rate should be modeled.
+Finally, the performance improvements of the data cache achieved by transferring data from the cache, and not from sites, can be simulated.
 \section{Conclusion}
@@
 Our first results are promising and suggest that this approach can be used to simulate load balancing strategies accurately.
-Regarding scalability we are confident that several computing centers can be simulated in a sufficient time, however this is part of future work.
+Regarding scalability, we are confident that several computing centers can be simulated in a sufficient time, however this is part of future work.
 

### 2018-10-29 paper/paper.tex
Modifications and comments RC

@@
 
-The amount of data processed by experiments in high energy physics (HEP) will drastically increase in the coming years due the upgrade of the LHC to the High-Luminosity Large Hadron Collider (HL-LHC), creating major challenges for the HEP community \cite{community}.
+The amount of data processed by experiments in high energy physics (HEP) will drastically increase in the coming years due to the upgrade of the LHC to the High-Luminosity Large Hadron Collider (HL-LHC), creating major challenges for the HEP community \cite{community}.
 \Cref{resources} shows the estimated required and provided CPU resources for the ATLAS experiment in the next ten years.
-Assuming a flat computing budget model, the estimated provided resources do not cover future needs for computing resource in the period from 2026 to 2028 due the start of the HL-LHC era.
-One of the key aspects to deal with this increasing load is the more efficient usage of existing resources.
+Assuming a flat computing budget model, the estimated provided resources do not cover future needs for computing resource in the period from 2026 to 2028 due to the start of the HL-LHC era.
+One of the key aspects to deal with this increasing load is the more efficient usage of existing and future resources.
 
-Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources of user jobs.
-Cloud resources, such as HNSciCloud and AWS, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs.
+Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources for user jobs.
+Cloud resources, such as the Helix Nebula Science Cloud and Amazon Web Services, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs.
 This results in a more complex scheduling, because each resource has different optimal job types and mixes.
-Furthermore, this leads to a more dynamic data placement, where fewer centers host data and thus remote access is increasing.
+Furthermore, a more dynamic data placement, where fewer centers host data, leads to an increase in the number of remote data accesses.
 
@@
 	\includegraphics[width=0.5\linewidth]{images/resources}
-	\caption[]{Estimated CPU resources of ATLAS \cite{community}}
+	\caption[]{Estimated CPU resource needs of the ATLAS experiment\cite{community}}
 	\label{resources}
@@
 We aim to create a simulation for the WLCG helping the HEP community to solve both challenges: a more efficient utilization of the grid and coping with the rising complexity of the system.
-Currently, there is no model or simulation in existence to evaluate the load balancing strategies of the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
+Currently, there is no model or simulation in existence to evaluate the load balancing strategies for the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
 \Cref{related} describes different attempts to simulate the WLCG or parts of it.
@@
 We chose Palladio to achieve a scalable simulation because it is a comparably abstract simulator (\Cref{related} gives an overview on other simulators and approaches).
-This leads to the research question of how we can use Palladio to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio.
-To answer this question, we created an extension for the simulator that enables us to simulate load balancing strategies.
+As the simulation of load balancing strategies is a new use case for Palladio, we created an extension for the simulator for this use case.
 Furthermore, we created an automated model parameter analysis and model creation in order to increase the usability of the simulation.
@@
 The starting point of the simulation process is matching the data sets from different data sources as a prerequisite  of the analysis.
-In our proof of concept we have three data sources: The CERN WMArchive, CMS GridKa job data and GridKa node performance data.
-These data sets have to be matched because each set contains different information about an executed job, but none contain the complete information about it.
+In our proof of concept we use data from three data sources: The CMS WMArchive, CMS GridKa job monitoring data and GridKa node performance data.
+These data sets have to be matched because each set contains different information about an executed job.
 
@@
 The usage model defines which and how many jobs are scheduled.
-To make sure that there is always a high load on the system, we used a closed workload with enough jobs that the system is fully utilized.
+To make sure that there is always a high load on the system, we simulate a closed workload with enough jobs that the system is fully utilized.
 Each job type has a configurable \textit{job share} of the overall load.
@@
 
-To obtain a comprehensive view both of the resource usage of CMS jobs executed at a WLCG grid site as well as the resources available there to process them, combining information from multiple data sources is required:
+To obtain a comprehensive view both of the resource usage of CMS jobs executed at a WLCG site as well as the resources available there to process them, combining information from multiple data sources is required:
 \begin{itemize}
-    \item We acquire historical job monitoring information from several global CMS monitoring systems: Job reports from JobMonitoring and Workflow and Data Management framework job reports (FWJR documents) are archived on the \emph{analytix} Hadoop cluster.
+    \item We acquire job monitoring information from several global CMS monitoring systems: Job reports from JobMonitoring and Workflow and Data Management framework job reports (FWJR documents) are archived on the \emph{analytix} Hadoop cluster.
     The latter are archived on the cluster as part of the WMArchive project \cite{WMArchiveWorkloadArchive2018}.
@@
 	\vspace{-10px}
-	\caption{Measured and simulated throughputs and CPU efficiency, job data from May 2018}
+	\caption{Measured and simulated throughputs and CPU efficiency using monitoring data from May 2018}
 	\label{fig:validation}
@@
 
-The CPU utilization in the simulation and a comparison with the measured CPU efficiency are shown in \Cref{fig:validation}, which also makes the initial ramp-up of the simulation visible.
+The CPU utilization in the simulation and a comparison with the measured average CPU efficiency are shown in \Cref{fig:validation}, which also makes the initial ramp-up of the simulation in the region up to 200000\,s visible.
 The overall average utilization in the simulation deviates from the reference by about 3\%, showing that even the current simplified model is capable of reproducing the metrics introduced above.
@@
 A first case study in the context of chemical computing showed that the Prediction Toolkit has high accuracy, but has not been tested on large systems yet \cite{rapidtesting}.
-However, due the required Runtime Toolkit CACTOS was not applicable in our use case.
+However, due to the required Runtime Toolkit CACTOS was not applicable in our use case.
 
@@
 
-We successfully simulated the CMS computing jobs at GridKa.
+As a proof of concept we successfully simulated CMS computing jobs at GridKa.
 To enable easy usage of the simulation we implemented an automated model parameter extraction and model creation.
+These methods can not only be used for CMS monitoring data, but are also applicable for computing jobs from any other experiment providing similar monitoring information.
 Furthermore, we extended Palladio to be able to simulate load balancing strategies.

### 2019-01-26 paper/paper.tex
Fit onto 8 pages again RC
Revise paper according to Reviews
- beware of the comma after the superscript numbers in the author list:
  you can remove it, just check the LaTeX template -> this is
  intentionally
- in the abstract, please prevent the line break between Tier and 1 (done)
- a reference to the GeliX Nebula Science Cloud project might be nice
(added https://www.hnscicloud.eu/ as footnote)
- "eclipse plugin": what is eclipse (and shoudn't it be capitalized)? ->
  Ecplise IDE
- "The MONARC project was started 1998" -> "in 1998"

@@
 
-	This paper presents a proof of concept in which the computing jobs at the Tier 1 center GridKa are modeled and simulated.
+	This paper presents a proof of concept in which the computing jobs at the \mbox{Tier 1} center GridKa are modeled and simulated.
 To model the computing jobs we extended the Palladio simulator with a mechanism to simulate load balancing strategies.
@@
 Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources for user jobs.
-Cloud resources, such as the Helix Nebula Science Cloud and Amazon Web Services, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs.
+Cloud resources, such as the Helix Nebula Science Cloud\footurl{http://www.helix-nebula.eu/} and Amazon Web Services, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs.
 This results in a more complex scheduling, because each resource has different optimal job types and mixes.
@@
 \label{sec:palladio}
-Palladio is a model driven software architecture simulator, which is implemented as an eclipse plugin and developed by Karlsruhe Institute of Technology (KIT), FZI Research Center for Information Technology, and University of Paderborn.
+Palladio is a model driven software architecture simulator, which is implemented as a plugin for the Eclipse IDE and developed by Karlsruhe Institute of Technology (KIT), FZI Research Center for Information Technology, and University of Paderborn.
 Palladio predicts quality of software properties (e.g.
-performance) using several models of a system \cite{BECKER20093}.
-These models are:
+performance) using the following models of a system \cite{BECKER20093}:
 \begin{itemize}
+	\vspace{-0.2pc}
+	\itemsep-0.2em
 	\item The component model: Specifies the structure and behavior of the components independently from their later usage. This allows reuse of the components, but requires its parametrization.
@@
 	\item The usage model: Defines the usage of a system regarding workload, user behavior, and parameters.
-
 \end{itemize}
+	\vspace{-0.2pc}
 Once these models are created, Palladio can be used to simulate the system choosing one of several supported simulators.
@@
 	\vspace{-10px}
-	\caption{Measured and simulated throughputs and CPU efficiency using monitoring data from May 2018}
+	\caption{Measured and simulated throughputs and CPU efficiency job data from May 2018}
 	\label{fig:validation}
@@
 \label{related}
-The MONARC project was started 1998 to help with the development of the initial design of the WLCG at CERN \cite{monarc2000models}.
+The MONARC project was started in 1998 to help with the development of the initial design of the WLCG at CERN \cite{monarc2000models}.
 However, after the initial design was developed, the project was discontinued in 2000 and thus no model and simulation of the WLCG was created. 

### 2019-01-28 paper/paper.tex
updated authors superscript, this is the version to be submitted.

@@
 \author{
-	\firstname{Ren\'e} \lastname{Caspart}\inst{1}\fnsep\thanks{\email{rene.caspart@cern.ch}} \and
-	\firstname{Patrick} \lastname{Firnkes}\inst{2}\fnsep\thanks{\email{patrick.firnkes@student.kit.edu}} \and
-	\firstname{Manuel} \lastname{Giffels}\inst{1}\fnsep\thanks{\email{manuel.giffels@cern.ch}} \and
-	\firstname{Anne} \lastname{Koziolek}\inst{2}\fnsep\thanks{\email{koziolek@kit.edu}} \and
-	\firstname{G\"unter} \lastname{Quast}\inst{1}\fnsep\thanks{\email{g.quast@kit.edu}} \and
-	\firstname{Ralf} \lastname{Reussner}\inst{2}\fnsep\thanks{\email{reussner@kit.edu}} \and
-	\firstname{Maximilian} \lastname{Stemmer-Grabow}\inst{2}\fnsep\thanks{\email{mail@mxsg.de}}
+	\firstname{Ren\'e} \lastname{Caspart}\inst{1}\thanks{\email{rene.caspart@cern.ch}} \and
+	\firstname{Patrick} \lastname{Firnkes}\inst{2}\thanks{\email{patrick.firnkes@student.kit.edu}} \and
+	\firstname{Manuel} \lastname{Giffels}\inst{1}\thanks{\email{manuel.giffels@cern.ch}} \and
+	\firstname{Anne} \lastname{Koziolek}\inst{2}\thanks{\email{koziolek@kit.edu}} \and
+	\firstname{G\"unter} \lastname{Quast}\inst{1}\thanks{\email{g.quast@kit.edu}} \and
+	\firstname{Ralf} \lastname{Reussner}\inst{2}\thanks{\email{reussner@kit.edu}} \and
+	\firstname{Maximilian} \lastname{Stemmer-Grabow}\inst{2}\thanks{\email{mail@mxsg.de}}
 }
